### 2024-02-13 Manuscript_1.Rmd
Figure 3 done - Polish Figure 4

---
title: "Manuscript_1"
author: Deborah Gérard^[University of Luxembourg - FSTM - DLSM - Systems Biology group - Epigenetics team]
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    keep_md: true
    df_print: paged
    toc: true
    toc_depth: 6
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      results = "as.is", 
                      fig.height = 3, 
                      cache = TRUE, 
                      tidy = TRUE, 
                      fig.pos = "H")
```

#### *1. Create a Singularity container*

Work within a Singularity container to ensure reproducibility

```{bash Sing_crea, eval = FALSE, echo = TRUE}
# Initiate virtual box Vagrant
cd $HOME/Manuscript_1_vagrant

vagrant up --provision
# vagrant reload
vagrant ssh

# Check the number of allocated cores and memory
free -h
nproc

# Check singularity version
singularity version   # 3.9.0

# Prepare the singularity definition file
mkdir Manuscript_1_singularity && cd Manuscript_1_singularity
touch Manuscript_1_singularity.def

# Build the temporary container based on Ubuntu 20.04 from Docker for development 
sudo singularity build --sandbox Manuscript_1_singularity_tmp Manuscript_1_singularity.def

# Run the container in writable mode to make changes
sudo singularity shell --writable Manuscript_1_singularity_tmp

# Now that all necessay packages and library have been installed, build the container
sudo singularity build Manuscript_1_singularity.sif Manuscript_1_singularity_tmp

# Start R
singularity exec Manuscript_1_singularity.sif R

# Check R version
R.Version()
```

**Note**: R version is 4.2.3 (2023-03-15)

##### *1.1 Customize the Singularity container*

Install different R libraries

```{r cust, eval = FALSE, echo = TRUE}
# tidyverse, ggpubr, rstatix
install.packages(c("tidyverse", 
                   "ggpubr", 
                   "rstatix"))
# plotgardener
BiocManager::install("plotgardener")

# BSgenome.Hsapiens.UCSC.hg38
BiocManager::install("BSgenome.Hsapiens.UCSC.hg38")

# extrafont
install.packages("extrafont")

# showtext
install.packages("showtext")

# TxDb.Hsapiens.UCSC.hg38.knownGene
BiocManager::install("TxDb.Hsapiens.UCSC.hg38.knownGene")

# org.Hs.eg.db
BiocManager::install("org.Hs.eg.db")

# JASPAR2020
BiocManager::install("JASPAR2020")

# TFBSTools
BiocManager::install("TFBSTools")

# ggseqlogo
install.packages("ggseqlogo")

# TxDb.Hsapiens.UCSC.hg38.refGene
BiocManager::install("TxDb.Hsapiens.UCSC.hg38.refGene")

# LDlinkR
install.packages("LDlinkR")

# AllelicImbalance
BiocManager::install("AllelicImbalance")

# ggflowchart
install.packages("ggflowchart")

# Issue between dplyr, BiocFilrCache and biomaRt - > downgrade dbplyr
install.packages("devtools")
devtools::install_version("dbplyr", version = "2.3.4")
```

Load libraries and check their versions

```{r load_lib, eval = FALSE, echo = TRUE}
# Load
library("tidyverse")
library("rstatix")
library("ggpubr")
library("plotgardener")
#library("BSgenome.Hsapiens.UCSC.hg38")
#library("TxDb.Hsapiens.UCSC.hg38.knownGene")
library("TxDb.Hsapiens.UCSC.hg38.refGene")
library("org.Hs.eg.db")
library("extrafont")
library("showtext")
library("JASPAR2020")
library("TFBSTools")
library("ggseqlogo")
library("biomaRt")
library("LDlinkR")
library("AllelicImbalance")
library("png")
library("AnnotationDbi")
library("scales")
library("flowCore")
library("ggcyto")

# Check
packageVersion("tidyverse") #2.0.0
packageVersion("rstatix") #0.7.2
packageVersion("ggpubr")  #0.6.0
packageVersion("plotgardener")  #1.4.2
#packageVersion("BSgenome.Hsapiens.UCSC.hg38") #1.4.5
#packageVersion("TxDb.Hsapiens.UCSC.hg38.knownGene") #3.16.0
packageVersion("TxDb.Hsapiens.UCSC.hg38.refGene") #3.15.0
packageVersion("org.Hs.eg.db")  #3.16.0
packageVersion("extrafont") #0.19
packageVersion("showtext")  #0.9.6
packageVersion("JASPAR2020")  #0.99.10
packageVersion("TFBSTools") #1.36.0
packageVersion("ggseqlogo") #0.1
packageVersion("biomaRt") #2.54.1
packageVersion("LDlinkR") #1.3.0
packageVersion("AllelicImbalance")  #1.36.0
packageVersion("png") #0.1.8
packageVersion("AnnotationDbi") #1.60.2
packageVersion("dbplyr")  #2.3.4

# Import all available fonts
font_import()
```

# FIGURES
### FIGURE 1
```{r fig1, eval = FALSE, echo = TRUE}
########## FIGURE 1 ##########
##############################
# Save as a PDF
pdf("/home/vagrant/Manuscript_1/FIGURE1/FIGURE1.pdf",
    width = 8.3, 
    height = 11.7)

# Create a A4 blank page
pageCreate(width = 8.3, 
           height = 11.7, 
           default.units = "inches",
           showGuides = FALSE)

#### PANEL A - 
# text Figure 1
plotText(label = "Figure 1", 
         fontsize = 14,
         x = 0.25, 
         y = 0.25, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")
# text A
plotText(label = "A", 
         fontsize = 12,
         x = 0.25, 
         y = 0.5, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# text B
plotText(label = "B", 
         fontsize = 12,
         x = 0.25, 
         y = 1.5, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# text C
plotText(label = "C", 
         fontsize = 12,
         x = 0.25, 
         y = 4.5, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# text D
plotText(label = "D", 
         fontsize = 12,
         x = 0.25, 
         y = 8.25, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

###################
#### Figure 1A ####
###################
Fig1A = readPNG("/home/vagrant/Manuscript_1/FIGURE1/Fig1A.png")

# Plot Figure 1A
plotRaster(image = Fig1A,
           x = 0.375,
           y = 0.625,
           width = 7.55,
           height = 1.0,
           just = c("left", 
                    "top"),
           interpolate = FALSE)

###################
#### Figure 1B ####
###################
# Load data obtained from Jochen
RNAseq_odd = read_delim("/home/vagrant/Manuscript_1/FIGURE1/REFORMAT_tpm_gene_sets.gsa.txt",
                        delim = "\t",
                        col_names = TRUE)

#RNAseq_odd = read_delim("/Volumes/deborah.gerard/Documents/Manuscript_1/FIGURE1/REFORMAT_tpm_gene_sets.gsa.txt",
#                        delim = "\t",
#                        col_names = TRUE)

# Calculate odd ratio
RNAseq_odd = RNAseq_odd %>% 
  dplyr::filter(!VARIABLE %in% c("microglia")) %>% 
  mutate(odd_R = exp(BETA)) %>% 
  mutate(VARIABLE = factor(VARIABLE, 
                           levels = c("smNPC",
                                      "D15negsort",
                                      "D15possort",
                                      "D30postsort",
                                      "D50negsort",
                                      "D50possort",
                                      "astrocytes")))

# Plot
fig1B = ggplot(RNAseq_odd,
       aes(x = odd_R,
           y = VARIABLE)) +
  geom_point(shape = 16, size = 3) +
  geom_errorbarh(aes(xmin = odd_R - SE, 
                     xmax = odd_R + SE), 
                 height = 0.25) +
  geom_vline(xintercept = 1.0, linetype = "dashed", color = "grey") +
  theme_classic() +
  theme(axis.title = element_text(face = "bold"),
        axis.text = element_text(face = "bold")) +
  scale_y_discrete(labels = c("smNPC" = "smNPC", 
                              "D15negsort" = "non-mDAN D15", 
                              "D15possort" = "mDAN D15", 
                              "D30postsort" = "mDAN D30", 
                              "D50negsort" = "non-mDAN D50",
                              "D50possort" = "mDAN D50",
                              "astrocytes" = "Astrocytes")) +
  geom_text(label = round(RNAseq_odd$P, 
                          digits = 3),
            nudge_y = 0.40) +
  xlab("Odds ratio (95% Confidence Interval)") +
  ylab("")

# Place
plotGG(plot = fig1B, 
       x = 0.375, 
       y = 1.75, 
       width = 3.0, 
       height = 2.5,
       just = c("left", "top"), 
       default.units = "inches")

###################
#### Figure 1C ####
###################
Fig1C_pip = readPNG("/home/vagrant/Manuscript_1/FIGURE1/Fig1C_pipeline.png")

Fig1C_tab1.1 = readPNG("/home/vagrant/Manuscript_1/FIGURE1/Fig1C_SNPs_tab1.png/Fig1C_SNPs_tab1.png-1.png")
Fig1C_tab1.2 = readPNG("/home/vagrant/Manuscript_1/FIGURE1/Fig1C_SNPs_tab1.png/Fig1C_SNPs_tab1.png-2.png")

# Plot Figure 1C
plotRaster(image = Fig1C_pip,
           x = 0.375,
           y = 5.25,
           width = 3.0,
           height = 2.5,
           just = c("left", 
                    "top"),
           interpolate = FALSE)

plotRaster(image = Fig1C_tab1.1,
           x = 3.375,
           y = 5.25,
           width = 2.3,
           height = 2.5,
           just = c("left", 
                    "top"),
           interpolate = FALSE)

plotRaster(image = Fig1C_tab1.2,
           x = 5.8,
           y = 5.25,
           width = 2.3,
           height = 2.5,
           just = c("left", 
                    "top"),
           interpolate = FALSE)

#diag_edge_data = tibble(from = c("~17x10^6\nPD GWAS SNPs",
#                                 "~3400 significant\nPD GWAS SNPs",
#                                 "28 significant\nPD rSNPs"),
#                        to = c("~3400 significant\nPD GWAS SNPs",
#                               "28 significant\nPD rSNPs",
#                               "3 significant\nPD rSNPs"))
#
#diag_node_data = tibble(name = c("~17x10^6\nPD GWAS SNPs",
#                                 "~3400 significant\nPD GWAS SNPs",
#                                 "28 significant\nPD rSNPs",
#                                "3 significant\nPD rSNPs"),
#                        label = name)

# Diagramme
# fig1C = ggflowchart(diag_edge_data,
#                    diag_node_data,
#                    text_size = 3.0,
#                    arrow_size = 0.2)

# Place the diagramme
#plotGG(plot = fig1C, 
#       x = 0.5, 
#       y = 4.5, 
#      width = 1.5, 
#       height = 3.0,
#       just = c("left", "top"), 
#      default.units = "inches")

###################
#### Figure 1D ####
###################

# Locus plot
# Load the Nalls et al. GWAS data
nalls_allSNPs.HG38.GR = readRDS(
  "/home/vagrant/Manuscript_1/FIGURE1/nalls_allSNPs.HG38.GR.rds")

#nalls_allSNPs.HG38.GR = readRDS(
#  "/Volumes/deborah.gerard/Documents/Manuscript_1/FIGURE1/nalls_allSNPs.HG38.GR.rds")

# Here our SNPs of interest and their location
snp = GRanges(seqnames = c("chr10", "chr4", "chr4"),
              ranges = IRanges(start = c(119651404, 987107, 76213716),
                               end = c(119651405, 987108, 76213717)))

# Retrieve them from Nalls et al data
snp_Nalls = subsetByOverlaps(nalls_allSNPs.HG38.GR, snp) %>% 
  as_tibble() %>% 
  dplyr::select(-width, -strand) %>% 
  mutate(rsID = c("rs144814361", "rs1465922", "rs11248061"))

# Visualise the genomic region per SNP (+- 400100 bp)
# rs144814361
r = 40100

# Filter for the SNP in BAG3 regulatory region
Nalls_sel_reg.BAG3 = nalls_allSNPs.HG38.GR %>% 
  as_tibble() %>% 
  dplyr::filter(seqnames == "chr10",
         between(base_pair_position, 119651405 - r, 119651405 + r)) %>% 
  mutate(minus_log_10_p = -log10(p),
         GWAS_p_sig = if_else(minus_log_10_p > -log10(5e-8), "yes", "no"),
         rsID = if_else(seqnames == "chr10" & base_pair_position == 119651405, "rs144814361", ""))

# Get SNPs info from Ensembl
snp.ensembl = useEnsembl(biomart = "snp", dataset = "hsapiens_snp")

# SNP info
query.snp = getBM(c("ensembl_gene_stable_id", 
                            "refsnp_id", 
                            "chr_name", 
                            "chrom_start", 
                            "chrom_end", 
                            "minor_allele", 
                            "minor_allele_freq"),
                          filters = "snp_filter",
                          values = c("rs144814361", "rs1465922", "rs11248061"),
                  mart = snp.ensembl)

# Get rsID of SNPs that are significant and append 
position_chr10 = Nalls_sel_reg.BAG3 %>% 
  mutate(chr = gsub("chr", "", seqnames)) %>% 
  unite(col = "position", c(chr, end, base_pair_position), sep = ":", remove = FALSE) %>% 
  filter(GWAS_p_sig != "no") %>%
  dplyr::select(position) %>% 
  pull()

rsID_sig_SNPs.chr10 = getBM(c("refsnp_id", "chr_name", "chrom_start", "chrom_end", "allele", "minor_allele"),
                      filters = "chromosomal_region",
                      values = position_chr10,
                      mart = snp.ensembl)

rsID_sig_SNPs.chr10 = rsID_sig_SNPs.chr10 %>% 
  unite(col = "position", chr_name:chrom_end, sep = ":", remove = FALSE)

Nalls_sel_reg.BAG3 = Nalls_sel_reg.BAG3 %>% 
  mutate(chr = gsub("chr", "", seqnames)) %>% 
  unite(col = "position", c(chr, end, base_pair_position),
        sep = ":", remove = FALSE) %>% 
  left_join(rsID_sig_SNPs.chr10, by = "position")

# Calculate R2 to check if SNPs are in LD
BAG3_SNPs_LD = Nalls_sel_reg.BAG3 %>% 
  dplyr::select(refsnp_id) %>% 
  drop_na()

BAG3_R2 = LDproxy(snp = "rs144814361",
                   pop = "EUR",
                   r2d = "r2",
                   token = "5bba40af90d7",
                   genome_build = "grch38_high_coverage")

BAG3_R2 = BAG3_R2 %>% 
  as_tibble() %>% 
  dplyr::select(RS_Number, Coord, R2) %>% 
  dplyr::rename(refsnp_id = RS_Number)

# Plot
BAG3.pl = Nalls_sel_reg.BAG3 %>% 
  unite("Coord", seqnames, base_pair_position, sep = ":", remove = FALSE) %>% 
  left_join(BAG3_R2, by = "Coord") %>% 
  dplyr::select(seqnames, base_pair_position, p, refsnp_id.x,R2) %>% 
  dplyr::rename(chrom = seqnames,
               pos = base_pair_position,
               snp = refsnp_id.x, 
               LD = R2) %>% 
  mutate(chrom = as.character(chrom)) %>% 
  group_by(LD) %>% 
  mutate(LD_grp = cut(LD, seq(0, 1.0, 0.2)))

BAG3.pl$LD_grp = addNA(BAG3.pl$LD_grp)

BAG3_leadSNP = "rs144814361"

BAG3_MP = plotManhattan(data = BAG3.pl,
              chrom = "chr10",
              chromstart = 119611405,
              chromend = 119691405,
              assembly = "hg38",
              fill = colorby("LD_grp",
                             palette = colorRampPalette(c("#35b779",
                                                          "#21918c", 
                                                          "#31688e",
                                                          "#443983", 
                                                          "#440154", 
                                                          "#CCCCCC"))),
        trans = "-log10", 
        sigLine = TRUE, col = "grey",
        lty = 2, range = c(0, 16),
        leadSNP = list(
        snp = BAG3_leadSNP,
        pch = 18,
        cex = 0.75,
        fill = "#990066",
        fontsize = 8),
              x = 0.5, 
              y = 8.5, 
              width = 1.5, 
              height = 2.5,
              just = c("left", "top"), 
              default.units = "inches")

## Annotate genome label
annoGenomeLabel(
    plot = BAG3_MP, x = 0.5, y = 11,
    fontsize = 8, scale = "Mb",
    just = c("left", "top"), default.units = "inches"
)

## Annotate y-axis
annoYaxis(
    plot = BAG3_MP,
    at = c(0, 2, 4, 6, 8, 10, 12),
    axisLine = TRUE, fontsize = 8
)

## Plot y-axis label
plotText(
    label = "-log10(p-value)", x = 0.125, y = 10.125, rot = 90,
    fontsize = 8, fontface = "bold", just = "center",
    default.units = "inches"
)

# Plot gene BAG3
plotGenes(
    chrom = "chr10", chromstart = 119611405, chromend = 119691405,
    x = 0.5, y = 8.5, width = 1.5, height = 0.5,
    just = c("left", "top"), default.units = "inches",
    fontcolor = "black",
    fill = "black",
    assembly = assembly(Genome = "hg38refGene", 
                      TxDb = "TxDb.Hsapiens.UCSC.hg38.refGene", 
                      OrgDb = "org.Hs.eg.db")
)

# Filter for the SNP in IDUA regulatory region
Nalls_sel_reg.IDUA = nalls_allSNPs.HG38.GR %>% 
  as_tibble() %>% 
  filter(seqnames == "chr4",
         between(base_pair_position, 987108 - r, 987108 + r)) %>% 
  mutate(minus_log_10_p = -log10(p),
         GWAS_p_sig = if_else(minus_log_10_p > -log10(5e-8), "yes", "no"),
         rsID = if_else(seqnames == "chr4" & base_pair_position == 987108, "rs11248061", ""))

# Get rsID of SNPs that significant and append 
position_chr4.IDUA = Nalls_sel_reg.IDUA %>% 
  mutate(chr = gsub("chr", "", seqnames)) %>% 
  unite(col = "position", c(chr, end, base_pair_position), sep = ":", remove = FALSE) %>% 
  filter(GWAS_p_sig != "no") %>%
  dplyr::select(position) %>% 
  pull()

rsID_sig_SNPs.chr4.IDUA = getBM(c("refsnp_id", "chr_name", "chrom_start", "chrom_end", "allele", "minor_allele",
                        "minor_allele_freq"),
                      filters = "chromosomal_region",
                      values = position_chr4.IDUA,
                      mart = snp.ensembl)

rsID_sig_SNPs.chr4.IDUA = rsID_sig_SNPs.chr4.IDUA %>% 
  unite(col = "position", chr_name:chrom_end, sep = ":", remove = FALSE)

Nalls_sel_reg.IDUA = Nalls_sel_reg.IDUA %>% 
  mutate(chr = gsub("chr", "", seqnames)) %>% 
  unite(col = "position", c(chr, end, base_pair_position),
        sep = ":", remove = FALSE) %>% 
  left_join(rsID_sig_SNPs.chr4.IDUA, by = "position")

# Calculate R2 to check if SNPs are in LD
IDUA_SNPs_LD = Nalls_sel_reg.IDUA %>% 
  dplyr::select(refsnp_id) %>% 
  drop_na()

IDUA_SNPs_LD = IDUA_SNPs_LD %>% 
  filter(!str_detect(refsnp_id, "^CR"))

IDUA_R2 = LDproxy(snp = "rs11248061",
                   pop = "EUR",
                   r2d = "r2",
                   token = "5bba40af90d7",
                   genome_build = "grch38_high_coverage")

IDUA_R2 = IDUA_R2 %>% 
  as_tibble() %>% 
  dplyr::select(RS_Number, Coord, R2) %>% 
  dplyr::rename(refsnp_id = RS_Number)

# Plot
IDUA.pl = Nalls_sel_reg.IDUA %>% 
  unite("Coord", seqnames, base_pair_position, sep = ":", remove = FALSE) %>% 
  distinct(Coord, .keep_all = TRUE) %>% 
  left_join(IDUA_R2, by = "Coord") %>% 
  dplyr::select(seqnames, base_pair_position, p, refsnp_id.x, R2) %>% 
  dplyr::rename(chrom = seqnames,
               pos = base_pair_position,
               snp = refsnp_id.x, 
               LD = R2) %>% 
  mutate(chrom = as.character(chrom)) %>% 
  group_by(LD) %>% 
  mutate(LD_grp = cut(LD, seq(0, 1.0, 0.2)))

IDUA.pl$LD_grp = addNA(IDUA.pl$LD_grp)

IDUA_leadSNP = "rs11248061"

IDUA_MP = plotManhattan(data = IDUA.pl,
              chrom = "chr4",
              chromstart = 947108,
              chromend = 1027108,
              assembly = "hg38",
              fill = colorby("LD_grp",
                             palette = colorRampPalette(c("#35b779",
                                                          "#21918c", 
                                                          "#31688e",
                                                          "#443983", 
                                                          "#440154", 
                                                          "#CCCCCC"))),
        trans = "-log10", 
        sigLine = TRUE, col = "grey",
        lty = 2, range = c(0, 16),
        leadSNP = list(
        snp = IDUA_leadSNP,
        pch = 18,
        cex = 0.75,
        fill = "#990066",
        fontsize = 8),
              x = 2.5, 
              y = 8.5, 
              width = 1.5, 
              height = 2.5,
              just = c("left", "top"), 
              default.units = "inches")

## Annotate genome label
annoGenomeLabel(
    plot = IDUA_MP, x = 2.5, y = 11,
    fontsize = 8, scale = "Mb",
    just = c("left", "top"), default.units = "inches"
)

## Annotate y-axis
annoYaxis(
    plot = IDUA_MP,
    at = c(0, 2, 4, 6, 8, 10, 12),
    axisLine = TRUE, fontsize = 8
)

## Plot y-axis label
plotText(
    label = "-log10(p-value)", x = 2.125, y = 10.125, rot = 90,
    fontsize = 8, fontface = "bold", just = "center",
    default.units = "inches"
)

# Plot gene IDUA
plotGenes(
    chrom = "chr4", chromstart = 947108, chromend = 1027108,
    x = 2.5, y = 8.5, width = 1.5, height = 0.5,
    just = c("left", "top"), default.units = "inches",
    fontcolor = "black",
    fill = "black",
    assembly = assembly(Genome = "hg38refGene", 
                      TxDb = "TxDb.Hsapiens.UCSC.hg38.refGene", 
                      OrgDb = "org.Hs.eg.db")
)

# Filter for the SNP in SCARB2 regulatory region
Nalls_sel_reg.SCARB2 = nalls_allSNPs.HG38.GR %>% 
  as_tibble() %>% 
  filter(seqnames == "chr4",
         between(base_pair_position, 76213717 - r, 76213717 + r)) %>% 
  mutate(minus_log_10_p = -log10(p),
         GWAS_p_sig = if_else(minus_log_10_p > -log10(5e-8), "yes", "no"),
         rsID = if_else(seqnames == "chr4" & base_pair_position == 76213717, "rs1465922", ""))

# Get rsID of SNPs that significant and append 
position_chr4.SCARB2 = Nalls_sel_reg.SCARB2 %>% 
  mutate(chr = gsub("chr", "", seqnames)) %>% 
  unite(col = "position", c(chr, end, base_pair_position), sep = ":", remove = FALSE) %>% 
  filter(GWAS_p_sig != "no") %>%
  dplyr::select(position) %>% 
  pull()

rsID_sig_SNPs.chr4.SCARB2 = getBM(c("refsnp_id", "chr_name", "chrom_start", "chrom_end", "allele", "minor_allele",
                        "minor_allele_freq"),
                      filters = "chromosomal_region",
                      values = position_chr4.SCARB2,
                      mart = snp.ensembl)

rsID_sig_SNPs.chr4.SCARB2 = rsID_sig_SNPs.chr4.SCARB2 %>% 
  unite(col = "position", chr_name:chrom_end, sep = ":", remove = FALSE)

Nalls_sel_reg.SCARB2 = Nalls_sel_reg.SCARB2 %>% 
  mutate(chr = gsub("chr", "", seqnames)) %>% 
  unite(col = "position", c(chr, end, base_pair_position),
        sep = ":", remove = FALSE) %>% 
  left_join(rsID_sig_SNPs.chr4.SCARB2, by = "position")

# Calculate R2 to check if SNPs are in LD
SCARB2_SNPs_LD = Nalls_sel_reg.SCARB2 %>% 
  dplyr::select(refsnp_id) %>% 
  drop_na()

SCARB2_R2 = LDproxy(snp = "rs1465922",
                   pop = "EUR",
                   r2d = "r2",
                   token = "5bba40af90d7",
                   genome_build = "grch38_high_coverage")

SCARB2_R2 = SCARB2_R2 %>% 
  as_tibble() %>% 
  dplyr::select(RS_Number, Coord, R2) %>% 
  dplyr::rename(refsnp_id = RS_Number)

# Plot
SCARB2.pl = Nalls_sel_reg.SCARB2 %>% 
  unite("Coord", seqnames, base_pair_position, sep = ":", remove = FALSE) %>% 
  left_join(SCARB2_R2, by = "Coord") %>% 
  dplyr::select(seqnames, base_pair_position, p, refsnp_id.x, R2) %>% 
  dplyr::rename(chrom = seqnames,
               pos = base_pair_position,
               snp = refsnp_id.x, 
               LD = R2) %>% 
  mutate(chrom = as.character(chrom)) %>% 
  group_by(LD) %>% 
  mutate(LD_grp = cut(LD, seq(0, 1.0, 0.2)))

SCARB2.pl$LD_grp = addNA(SCARB2.pl$LD_grp)

SCARB2_leadSNP = "rs1465922"

SCARB2_MP = plotManhattan(data = SCARB2.pl,
              chrom = "chr4",
              chromstart = 76173717,
              chromend = 76253717,
              assembly = "hg38",
              fill = colorby("LD_grp",
                             palette = colorRampPalette(c("#35b779",
                                                          "#21918c", 
                                                          "#31688e",
                                                          "#443983", 
                                                          "#440154", 
                                                          "#CCCCCC"))),
        trans = "-log10", 
        sigLine = TRUE, col = "grey",
        lty = 2, range = c(0, 16),
        leadSNP = list(
        snp = SCARB2_leadSNP,
        pch = 18,
        cex = 0.75,
        fill = "#990066",
        fontsize = 8),
              x = 4.5, 
              y = 8.5, 
              width = 1.5, 
              height = 2.5,
              just = c("left", "top"), 
              default.units = "inches")

## Annotate genome label
annoGenomeLabel(
    plot = SCARB2_MP, x = 4.5, y = 11,
    fontsize = 8, scale = "Mb",
    just = c("left", "top"), default.units = "inches"
)

## Annotate y-axis
annoYaxis(
    plot = SCARB2_MP,
    at = c(0, 2, 4, 6, 8, 10, 12),
    axisLine = TRUE, fontsize = 8
)

## Plot y-axis label
plotText(
    label = "-log10(p-value)", x = 4.125, y = 10.125, rot = 90,
    fontsize = 8, fontface = "bold", just = "center",
    default.units = "inches"
)

# Plot gene SCARB2
plotGenes(
    chrom = "chr4", chromstart = 76173717, chromend = 76253717,
    x = 4.5, y = 8.5, width = 1.5, height = 0.5,
    just = c("left", "top"), default.units = "inches",
    fontcolor = "black",
    fill = "black",
    assembly = assembly(Genome = "hg38refGene", 
                      TxDb = "TxDb.Hsapiens.UCSC.hg38.refGene", 
                      OrgDb = "org.Hs.eg.db")
)

# Legend LD score
plotLegend(
  legend = c("LD reference SNP",
             paste("0", "<", "r^2", "<= 0.2"),
             paste("0.2", "<", "r^2", "<= 0.4"),
             paste("0.4", "<", "r^2", "<= 0.6"),
             paste("0.6", "<", "r^2", "<= 0.8"),
             paste("0.8", "<", "r^2", "<= 1.0"),
             "no LD data"),
    fill = c("#990066", "#35b779", "#21918c", "#31688e", "#443983", "#440154", "#CCCCCC"), 
  cex = 0.75,
    pch = c(18, 19, 19, 19, 19, 19, 19), border = FALSE, x = 6.5, y = 9.25,
    width = 1.0, height = 1.0, just = c("left", "top"),
    default.units = "inches"
)

dev.off()
```

### FIGURE 2
Create the page layout that will contain all the necessary plots

```{r fig2, eval = FALSE, echo = TRUE}
########## FIGURE 2 ##########
##############################

# Save as a PDF
pdf("/home/vagrant/Manuscript_1/FIGURE2/FIGURE2.pdf",
    width = 8.3, 
    height = 11.7)

# Create a A4 blank page
pageCreate(width = 8.3, 
           height = 11.7, 
           default.units = "inches",
           showGuides = FALSE)

#### PANEL A - BAG3 locus
# text Figure 2
plotText(label = "Figure 2", 
         fontsize = 14,
         x = 0.25, 
         y = 0.25, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")
# text A
plotText(label = "A", 
         fontsize = 12,
         x = 0.25, 
         y = 0.5, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# ATACseq signal BAG3
# path where bigwig files are
bw.path = "/home/vagrant/epifunc/"

# 1st bio replicate smNPC
bw.smNPC_1 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/smNPC_I.bw"),
                        chrom = "chr10",
                        chromstart = 119611305,
                        chromend = 119691505)

# 2nd bio replicate smNPC
bw.smNPC_2 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/smNPC_II.bw"),
                        chrom = "chr10",
                        chromstart = 119611305,
                        chromend = 119691505)

# 3rd bio replicate smNPC
bw.smNPC_3 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/smNPC_III.bw"),
                        chrom = "chr10",
                        chromstart = 119611305,
                        chromend = 119691505)

# 1st bio replicate TH positive neurons D15
bw.TH.pos_D15_1 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D15_POS_I.bw"),
                        chrom = "chr10",
                        chromstart = 119611305,
                        chromend = 119691505)

# 2nd bio replicate TH positive neurons D15
bw.TH.pos_D15_2 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D15_POS_II.bw"),
                        chrom = "chr10",
                        chromstart = 119611305,
                        chromend = 119691505)

# 3rd bio replicate TH positive neurons D15
bw.TH.pos_D15_3 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D15_POS_III.bw"),
                        chrom = "chr10",
                        chromstart = 119611305,
                        chromend = 119691505)

# 1st bio replicate TH positive neurons D30
bw.TH.pos_D30_1 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D30_POS_I.bw"),
                        chrom = "chr10",
                        chromstart = 119611305,
                        chromend = 119691505)

# 2nd bio replicate TH positive neurons D30
bw.TH.pos_D30_2 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D30_POS_II.bw"),
                        chrom = "chr10",
                        chromstart = 119611305,
                        chromend = 119691505)

# 3rd bio replicate TH positive neurons D30
bw.TH.pos_D30_3 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D30_POS_III.bw"),
                        chrom = "chr10",
                        chromstart = 119611305,
                        chromend = 119691505)

# 1st bio replicate TH positive neurons D50
bw.TH.pos_D50_1 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/HFFTHmCherry_D50_possort_S1.bw"),
                        chrom = "chr10",
                        chromstart = 119611305,
                        chromend = 119691505)

# 2nd bio replicate TH positive neurons D50
bw.TH.pos_D50_2 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/HFFTHmCherry_D50_possort_S3.bw"),
                        chrom = "chr10",
                        chromstart = 119611305,
                        chromend = 119691505)

# 3rd bio replicate TH positive neurons D50
bw.TH.pos_D50_3 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/HFFTHmCherry_D50_possort_S4.bw"),
                        chrom = "chr10",
                        chromstart = 119611305,
                        chromend = 119691505)

# Add a scale next to the bigwig files - check the maximum to choose
scale.BAG3.max = max(c(bw.smNPC_1$score,
                       bw.smNPC_2$score,
                       bw.smNPC_3$score,
                       bw.TH.pos_D15_1$score,
                       bw.TH.pos_D15_2$score,
                       bw.TH.pos_D15_3$score,
                       bw.TH.pos_D30_1$score,
                       bw.TH.pos_D30_2$score,
                       bw.TH.pos_D30_3$score,
                       bw.TH.pos_D50_1$score,
                       bw.TH.pos_D50_2$score,
                       bw.TH.pos_D50_3$score))

print(scale.BAG3.max)

# Define parameters for the regions
region.p.BAG3 = pgParams(chrom = "chr10",
                         chromstart = 119611405,
                         chromend = 119691405,
                         assembly = "hg38",
                         range = c(0, scale.BAG3.max))

# Add the genomic label
plotGenomeLabel(
  chrom = "chr10",
  chromstart = 119611405,
  chromend = 119691405,
  assembly = "hg38",
  x = 0.75,
  y = 0.625,
  length = 4.0,
  default.units = "inches",
  scale = "Mb")

# ATACseq signal smNPC_I and scale
ATAC_smNPC_I = plotSignal(
  data = bw.smNPC_1, 
  params = region.p.BAG3,
  fill = "#D3D3D3", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 0.875, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_smNPC_I, 
          at = c(0, scale.BAG3.max),
          fontsize = 6)

# ATACseq signal smNPC_II
ATAC_smNPC_II = plotSignal(
  data = bw.smNPC_2, 
  params = region.p.BAG3,
  fill = "#C0C0C0", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 0.875, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_smNPC_II, 
          at = c(0, scale.BAG3.max),
          fontsize = 6)

# ATACseq signal smNPC_III
ATAC_smNPC_III = plotSignal(
  data = bw.smNPC_3, 
  params = region.p.BAG3,
  fill = "#A9A9A9", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 0.875, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_smNPC_III, 
          at = c(0, scale.BAG3.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D15 I
ATAC_TH.pos_D15_I = plotSignal(
  data = bw.TH.pos_D15_1, 
  params = region.p.BAG3,
  fill = "#B22222", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 1.25, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D15_I, 
          at = c(0, scale.BAG3.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D15 II
ATAC_TH.pos_D15_II = plotSignal(
  data = bw.TH.pos_D15_2, 
  params = region.p.BAG3,
  fill = "#A52A2A", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 1.25, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D15_II, 
          at = c(0, scale.BAG3.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D15 III
ATAC_TH.pos_D15_III = plotSignal(
  data = bw.TH.pos_D15_3, 
  params = region.p.BAG3,
  fill = "#8B0000", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 1.25, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D15_III, 
          at = c(0, scale.BAG3.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D30 I
ATAC_TH.pos_D30_I = plotSignal(
  data = bw.TH.pos_D30_1, 
  params = region.p.BAG3,
  fill = "#FF6347", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 1.625, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D30_I, 
          at = c(0, scale.BAG3.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D30 II
ATAC_TH.pos_D30_II = plotSignal(
  data = bw.TH.pos_D30_2, 
  params = region.p.BAG3,
  fill = "#FF0000", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 1.625, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D30_II, 
          at = c(0, scale.BAG3.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D30 III
ATAC_TH.pos_D30_III = plotSignal(
  data = bw.TH.pos_D30_3, 
  params = region.p.BAG3,
  fill = "#DC143C", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 1.625, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D30_III, 
          at = c(0, scale.BAG3.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D50 I
ATAC_TH.pos_D50_I = plotSignal(
  data = bw.TH.pos_D50_1, 
  params = region.p.BAG3,
  fill = "#FFA07A", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 2.0, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D50_I, 
          at = c(0, scale.BAG3.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D50 II
ATAC_TH.pos_D50_II = plotSignal(
  data = bw.TH.pos_D50_2, 
  params = region.p.BAG3,
  fill = "#FA8072", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 2.0, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D50_II, 
          at = c(0, scale.BAG3.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D50 III
ATAC_TH.pos_D50_III = plotSignal(
  data = bw.TH.pos_D50_3, 
  params = region.p.BAG3,
  fill = "#E9967A", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 2.0, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D50_III, 
          at = c(0, scale.BAG3.max),
          fontsize = 6)

# Gene track
plotGenes(chrom = "chr10",
  chromstart = 119611405,
  chromend = 119691405,
  assembly = assembly(Genome = "hg38refGene", 
                      TxDb = "TxDb.Hsapiens.UCSC.hg38.refGene", 
                      OrgDb = "org.Hs.eg.db"),
  fontcolor = "black",
  fill = "black",
  x = 0.75,
  y = 2.375,
  width = 4.0,
  height = 0.375,
  just = c("left", "top"),
  default.units = "inches")

# Add the samples name
plotText(label = "smNPC", 
         fontsize = 6, 
         fontcolor = "#A9A9A9",
         x = 0.75, 
         y = 0.875, 
         just = c("left", "top"),
         default.units = "inches")

plotText(label = "mDAN D15", 
         fontsize = 6, 
         fontcolor = "#B22222",
         x = 0.75, 
         y = 1.25, 
         just = c("left", "top"),
         default.units = "inches")

plotText(label = "mDAN D30", 
         fontsize = 6, 
         fontcolor = "#DC143C",
         x = 0.75, 
         y = 1.625, 
         just = c("left", "top"),
         default.units = "inches")

plotText(label = "mDAN D50", 
         fontsize = 6, 
         fontcolor = "#E9967A",
         x = 0.75, 
         y = 2.0, 
         just = c("left", "top"),
         default.units = "inches")

#### PANEL B - BAG3 and LHX1 expression (RNAseq)
# text B
plotText(label = "B", 
         fontsize = 12,
         x = 5.0, 
         y = 0.5, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# Load RPKM matrix obtained from Borja and rename columns
# dat.RPKM = read_delim("/home/vagrant/epifunc/RNAseq/RPKM_genename",
#                       delim = "\t", 
#                       col_names = TRUE)
# 
# dat.RPKM = dat.RPKM %>% 
#   dplyr::select(gene_id, gene_name, everything())

# RPKM data - Remove the negative sorted neurons and astrocytes
# dat.RPKM.filt = dat.RPKM %>% 
#   dplyr::select(!contains("negsort"), -starts_with("ASTRO")) %>% 
#   pivot_longer(cols = ends_with("bam"), names_to = "Sample", values_to = "RPKM") %>% 
#   mutate(Cond = case_when(str_detect(Sample, "D15_possort") ~ "Positively sorted neurons D15",
#                          str_detect(Sample, "D30_possort") ~ "Positively sorted neurons D30",
#                          str_detect(Sample, "D50_possort") ~ "Positively sorted neurons D50",
#                          TRUE ~ "smNPCs"),
#          Cond = factor(Cond, levels = c("smNPCs", "Positively sorted neurons D15",
#                                   "Positively sorted neurons D30", 
#                                   "Positively sorted neurons D50")),
#          gene_id = gsub("\\..*", "", gene_id))

# Save the matrix of expression
# write_rds(dat.RPKM.filt,
#           "/home/vagrant/Manuscript_1/FIGURE2/mat_RPKM.rds")

# Load it for making the plots
dat.RPKM.filt = read_rds("/home/vagrant/Manuscript_1/FIGURE2/mat_RPKM.rds")

# Expression for BAG3 and LHX1
BAG3.LHX1.exp = ggboxplot(dat.RPKM.filt %>% 
                            dplyr::filter(gene_name %in% c("BAG3", "LHX1")),
                          x = "Cond", 
                          y = "RPKM", 
                          add = "jitter", 
                          fill = "Cond", 
                          palette = c("#A9A9A9", "#B22222", "#DC143C", "#E9967A"),
                          facet.by = "gene_name",
                          xlab = "", 
                          ylab = "Reads Per Kilobase Million (RPKM)") +
  scale_x_discrete(labels = c("smNPC",
                              expression("mDAN D15"),
                              expression("mDAN D30"),
                              expression("mDAN D50"))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))

# Place the plot
plotGG(plot = BAG3.LHX1.exp, 
       x = 5.25, 
       y = 0.625, 
       width = 2, 
       height = 3.5,
       just = c("left", "top"), 
       default.units = "inches")

# Add the transcription factor binding motif (TFBS) for LHX1
pfm.LHX1 = getMatrixByID(JASPAR2020, 
                         ID = "MA1518.1")
LHX1 = new.env()

LHX1$LHX1 = pfm.LHX1@profileMatrix

LHX1 = as.list(LHX1)
LHX1_TFBS = ggseqlogo(LHX1)

# Place the plot
plotGG(plot = LHX1_TFBS, 
       x = 1.7125, 
       y = 2.625, 
       width = 1.5, 
       height = 1.5,
       just = c("left", "top"), 
       default.units = "inches")

# Add the SNP position that zoom in the TFBS
SNP_param = pgParams(chrom = "chr10",
                     chromstart = 119651405,
                     chromend = 119651405,
                     assembly = "hg38")

annoHighlight(plot = ATAC_smNPC_I, 
              params = SNP_param,
              fill = "#404788FF",
              y = 0.875, 
              height = 1.75, 
              just = c("left", "top"), 
              default.units = "inches")

annoZoomLines(plot = ATAC_smNPC_I, 
              params = SNP_param,
              y0 = 2.625, 
              x1 = c(2.25, 3.25), 
              y1 = 2.75, 
              default.units = "inches")

# text C
plotText(label = "C", 
         fontsize = 12,
         x = 0.25, 
         y = 4.25, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# ATACseq signal IDUA
# 1st bio replicate smNPC
bw.IDUA.smNPC_1 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/smNPC_I.bw"),
                        chrom = "chr4",
                        chromstart = 947008,
                        chromend = 1027208)

# 2nd bio replicate smNPC
bw.IDUA.smNPC_2 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/smNPC_II.bw"),
                        chrom = "chr4",
                        chromstart = 947008,
                        chromend = 1027208)

# 3rd bio replicate smNPC
bw.IDUA.smNPC_3 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/smNPC_III.bw"),
                        chrom = "chr4",
                        chromstart = 947008,
                        chromend = 1027208)

# 1st bio replicate TH positive neurons D15
bw.IDUA.TH.pos_D15_1 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D15_POS_I.bw"),
                        chrom = "chr4",
                        chromstart = 947008,
                        chromend = 1027208)

# 2nd bio replicate TH positive neurons D15
bw.IDUA.TH.pos_D15_2 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D15_POS_II.bw"),
                        chrom = "chr4",
                        chromstart = 947008,
                        chromend = 1027208)

# 3rd bio replicate TH positive neurons D15
bw.IDUA.TH.pos_D15_3 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D15_POS_III.bw"),
                        chrom = "chr4",
                        chromstart = 947008,
                        chromend = 1027208)

# 1st bio replicate TH positive neurons D30
bw.IDUA.TH.pos_D30_1 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D30_POS_I.bw"),
                        chrom = "chr4",
                        chromstart = 947008,
                        chromend = 1027208)

# 2nd bio replicate TH positive neurons D30
bw.IDUA.TH.pos_D30_2 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D30_POS_II.bw"),
                        chrom = "chr4",
                        chromstart = 947008,
                        chromend = 1027208)

# 3rd bio replicate TH positive neurons D30
bw.IDUA.TH.pos_D30_3 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D30_POS_III.bw"),
                        chrom = "chr4",
                        chromstart = 947008,
                        chromend = 1027208)

# 1st bio replicate TH positive neurons D50
bw.IDUA.TH.pos_D50_1 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/HFFTHmCherry_D50_possort_S1.bw"),
                        chrom = "chr4",
                        chromstart = 947008,
                        chromend = 1027208)

# 2nd bio replicate TH positive neurons D50
bw.IDUA.TH.pos_D50_2 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/HFFTHmCherry_D50_possort_S3.bw"),
                        chrom = "chr4",
                        chromstart = 947008,
                        chromend = 1027208)

# 3rd bio replicate TH positive neurons D50
bw.IDUA.TH.pos_D50_3 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/HFFTHmCherry_D50_possort_S4.bw"),
                        chrom = "chr4",
                        chromstart = 947008,
                        chromend = 1027208)

# Add a scale next to the bigwig files - check the maximum to choose
scale.IDUA.max = max(c(bw.IDUA.smNPC_1$score,
                       bw.IDUA.smNPC_2$score,
                       bw.IDUA.smNPC_3$score,
                       bw.IDUA.TH.pos_D15_1$score,
                       bw.IDUA.TH.pos_D15_2$score,
                       bw.IDUA.TH.pos_D15_3$score,
                       bw.IDUA.TH.pos_D30_1$score,
                       bw.IDUA.TH.pos_D30_2$score,
                       bw.IDUA.TH.pos_D30_3$score,
                       bw.IDUA.TH.pos_D50_1$score,
                       bw.IDUA.TH.pos_D50_2$score,
                       bw.IDUA.TH.pos_D50_3$score))

print(scale.IDUA.max)

# Define parameters for the regions
region.p.IDUA = pgParams(chrom = "chr4",
                         chromstart = 947108,
                         chromend = 1027108,
                         assembly = "hg38",
                         range = c(0, scale.IDUA.max))

# Add the genomic label
plotGenomeLabel(
  chrom = "chr4",
  chromstart = 947108,
  chromend = 1027108,
  assembly = "hg38",
  x = 0.75,
  y = 4.375,
  length = 4.0,
  default.units = "inches",
  scale = "Mb")

# ATACseq signal smNPC_I
ATAC_smNPC_I_IDUA = plotSignal(
  data = bw.IDUA.smNPC_1, 
  params = region.p.IDUA,
  fill = "#D3D3D3", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 4.625, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_smNPC_I_IDUA, 
          at = c(0, scale.IDUA.max),
          fontsize = 6)

# ATACseq signal smNPC_II
ATAC_smNPC_II_IDUA = plotSignal(
  data = bw.IDUA.smNPC_2, 
  params = region.p.IDUA,
  fill = "#C0C0C0", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 4.625, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_smNPC_II_IDUA, 
          at = c(0, scale.IDUA.max),
          fontsize = 6)

# Add the SNP position that zoom in the TFBS
SNP_param_IDUA = pgParams(chrom = "chr4",
                     chromstart = 987108,
                     chromend = 987108,
                     assembly = "hg38")

annoHighlight(plot = ATAC_smNPC_I_IDUA, 
              params = SNP_param_IDUA,
              fill = "#404788FF",
              y = 4.75, 
              height = 1.875, 
              just = c("left", "top"), 
              default.units = "inches")

annoZoomLines(plot = ATAC_smNPC_I_IDUA, 
              params = SNP_param_IDUA,
              y0 = 6.625, 
              x1 = c(2.25, 3.25), 
              y1 = 6.75, 
              default.units = "inches")

# ATACseq signal smNPC_III
ATAC_smNPC_III_IDUA = plotSignal(
  data = bw.IDUA.smNPC_3, 
  params = region.p.IDUA,
  fill = "#A9A9A9", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 4.625, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_smNPC_III_IDUA, 
          at = c(0, scale.IDUA.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D15 I
ATAC_TH.pos_D15_I_IDUA = plotSignal(
  data = bw.IDUA.TH.pos_D15_1, 
  params = region.p.IDUA,
  fill = "#B22222", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 5.0, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D15_I_IDUA, 
          at = c(0, scale.IDUA.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D15 II
ATAC_TH.pos_D15_II_IDUA = plotSignal(
  data = bw.IDUA.TH.pos_D15_2, 
  params = region.p.IDUA,
  fill = "#A52A2A", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 5.0, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D15_II_IDUA, 
          at = c(0, scale.IDUA.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D15 III
ATAC_TH.pos_D15_III_IDUA = plotSignal(
  data = bw.IDUA.TH.pos_D15_3, 
  params = region.p.IDUA,
  fill = "#8B0000", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 5.0, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D15_III_IDUA, 
          at = c(0, scale.IDUA.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D30 I
ATAC_TH.pos_D30_I_IDUA = plotSignal(
  data = bw.IDUA.TH.pos_D30_1, 
  params = region.p.IDUA,
  fill = "#FF6347", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 5.375, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D30_I_IDUA, 
          at = c(0, scale.IDUA.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D30 II
ATAC_TH.pos_D30_II_IDUA = plotSignal(
  data = bw.IDUA.TH.pos_D30_2, 
  params = region.p.IDUA,
  fill = "#FF0000", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 5.375, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D30_II_IDUA, 
          at = c(0, scale.IDUA.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D30 III
ATAC_TH.pos_D30_III_IDUA = plotSignal(
  data = bw.IDUA.TH.pos_D30_3, 
  params = region.p.IDUA,
  fill = "#DC143C", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 5.375, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D30_III_IDUA, 
          at = c(0, scale.IDUA.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D50 I
ATAC_TH.pos_D50_I_IDUA = plotSignal(
  data = bw.IDUA.TH.pos_D50_1, 
  params = region.p.IDUA,
  fill = "#FFA07A", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 5.75, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D50_I_IDUA, 
          at = c(0, scale.IDUA.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D50 II
ATAC_TH.pos_D50_II_IDUA = plotSignal(
  data = bw.IDUA.TH.pos_D50_2, 
  params = region.p.IDUA,
  fill = "#FA8072", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 5.75, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D50_II_IDUA, 
          at = c(0, scale.IDUA.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D50 III
ATAC_TH.pos_D50_III_IDUA = plotSignal(
  data = bw.IDUA.TH.pos_D50_3, 
  params = region.p.IDUA,
  fill = "#E9967A", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 5.75, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D50_III_IDUA, 
          at = c(0, scale.IDUA.max),
          fontsize = 6)

# Gene track
plotGenes(chrom = "chr4",
  chromstart = 947108,
  chromend = 1027108,
  assembly = assembly(Genome = "hg38refGene", 
                      TxDb = "TxDb.Hsapiens.UCSC.hg38.refGene", 
                      OrgDb = "org.Hs.eg.db"),
  fontcolor = "black",
  fill = "black",
  x = 0.75,
  y = 6.125,
  width = 4.0,
  height = 0.375,
  just = c("left", "top"),
  default.units = "inches")

# Add the samples name
plotText(label = "smNPC", 
         fontsize = 6, 
         fontcolor = "#A9A9A9",
         x = 0.75, 
         y = 4.625, 
         just = c("left", "top"),
         default.units = "inches")

plotText(label = "mDAN D15", 
         fontsize = 6, 
         fontcolor = "#B22222",
         x = 0.75, 
         y = 5.0, 
         just = c("left", "top"),
         default.units = "inches")

plotText(label = "mDAN D30", 
         fontsize = 6, 
         fontcolor = "#DC143C",
         x = 0.75, 
         y = 5.375, 
         just = c("left", "top"),
         default.units = "inches")

plotText(label = "mDAN D50", 
         fontsize = 6, 
         fontcolor = "#E9967A",
         x = 0.75, 
         y = 5.75, 
         just = c("left", "top"),
         default.units = "inches")

#### PANEL D - IDUA and ZBTB14 expression (RNAseq)
# text D
plotText(label = "D", 
         fontsize = 12,
         x = 5.0, 
         y = 4.25, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

IDUA.ZBTB14.exp = ggboxplot(dat.RPKM.filt %>% 
                            dplyr::filter(gene_name %in% c("IDUA", "ZBTB14", "SLC26A1")) %>% 
                              mutate(gene_name = factor(gene_name,
                                                        levels = c("IDUA", "ZBTB14", "SLC26A1"))),
                          x = "Cond", 
                          y = "RPKM", 
                          add = "jitter", 
                          fill = "Cond", 
                          palette = c("#A9A9A9", "#B22222", "#DC143C", "#E9967A"),
                          facet.by = "gene_name",
                          xlab = "", 
                          ylab = "Reads Per Kilobase Million (RPKM)") +
  scale_x_discrete(labels = c("smNPC",
                              "mDAN D15",
                              "mDAN D30",
                              "mDAN D50")) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))

# Place the plot
plotGG(plot = IDUA.ZBTB14.exp, 
       x = 5.25, 
       y = 4.25, 
       width = 3, 
       height = 3.5,
       just = c("left", "top"), 
       default.units = "inches")

# Add the transcription factor binding motif (TFBS) for ZBTB14
pfm.ZBTB14 = getMatrixByID(JASPAR2020, 
                         ID = "MA1650.1")
ZBTB14 = new.env()

ZBTB14$ZBTB14 = pfm.ZBTB14@profileMatrix

ZBTB14 = as.list(ZBTB14)
ZBTB14_TFBS = ggseqlogo(ZBTB14)

# Place the plot
plotGG(plot = ZBTB14_TFBS, 
       x = 1.7125, 
       y = 6.625, 
       width = 1.5, 
       height = 1.5,
       just = c("left", "top"), 
       default.units = "inches")

# text E
plotText(label = "E", 
         fontsize = 12,
         x = 0.25, 
         y = 8.0, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# ATACseq signal SCARB2
# 1st bio replicate smNPC
bw.SCARB2.smNPC_1 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/smNPC_I.bw"),
                        chrom = "chr4",
                        chromstart = 76173617,
                        chromend = 76253817)

# 2nd bio replicate smNPC
bw.SCARB2.smNPC_2 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/smNPC_II.bw"),
                        chrom = "chr4",
                        chromstart = 76173617,
                        chromend = 76253817)

# 3rd bio replicate smNPC
bw.SCARB2.smNPC_3 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/smNPC_III.bw"),
                        chrom = "chr4",
                        chromstart = 76173617,
                        chromend = 76253817)

# 1st bio replicate TH positive neurons D15
bw.SCARB2.TH.pos_D15_1 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D15_POS_I.bw"),
                        chrom = "chr4",
                        chromstart = 76173617,
                        chromend = 76253817)

# 2nd bio replicate TH positive neurons D15
bw.SCARB2.TH.pos_D15_2 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D15_POS_II.bw"),
                        chrom = "chr4",
                        chromstart = 76173617,
                        chromend = 76253817)

# 3rd bio replicate TH positive neurons D15
bw.SCARB2.TH.pos_D15_3 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D15_POS_III.bw"),
                        chrom = "chr4",
                        chromstart = 76173617,
                        chromend = 76253817)

# 1st bio replicate TH positive neurons D30
bw.SCARB2.TH.pos_D30_1 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D30_POS_I.bw"),
                        chrom = "chr4",
                        chromstart = 76173617,
                        chromend = 76253817)

# 2nd bio replicate TH positive neurons D30
bw.SCARB2.TH.pos_D30_2 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D30_POS_II.bw"),
                        chrom = "chr4",
                        chromstart = 76173617,
                        chromend = 76253817)

# 3rd bio replicate TH positive neurons D30
bw.SCARB2.TH.pos_D30_3 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/D30_POS_III.bw"),
                        chrom = "chr4",
                        chromstart = 76173617,
                        chromend = 76253817)

# 1st bio replicate TH positive neurons D50
bw.SCARB2.TH.pos_D50_1 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/HFFTHmCherry_D50_possort_S1.bw"),
                        chrom = "chr4",
                        chromstart = 76173617,
                        chromend = 76253817)

# 2nd bio replicate TH positive neurons D50
bw.SCARB2.TH.pos_D50_2 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/HFFTHmCherry_D50_possort_S3.bw"),
                        chrom = "chr4",
                        chromstart = 76173617,
                        chromend = 76253817)

# 3rd bio replicate TH positive neurons D50
bw.SCARB2.TH.pos_D50_3 = readBigwig(file = paste0(bw.path, 
                                      "BIGWIG/HFFTHmCherry_D50_possort_S4.bw"),
                        chrom = "chr4",
                        chromstart = 76173617,
                        chromend = 76253817)

# Add a scale next to the bigwig files - check the maximum to choose
scale.SCARB2.max = max(c(bw.SCARB2.smNPC_1$score,
                       bw.SCARB2.smNPC_2$score,
                       bw.SCARB2.smNPC_3$score,
                       bw.SCARB2.TH.pos_D15_1$score,
                       bw.SCARB2.TH.pos_D15_2$score,
                       bw.SCARB2.TH.pos_D15_3$score,
                       bw.SCARB2.TH.pos_D30_1$score,
                       bw.SCARB2.TH.pos_D30_2$score,
                       bw.SCARB2.TH.pos_D30_3$score,
                       bw.SCARB2.TH.pos_D50_1$score,
                       bw.SCARB2.TH.pos_D50_2$score,
                       bw.SCARB2.TH.pos_D50_3$score))

print(scale.SCARB2.max)

# Define parameters for the regions
region.p.SCARB2 = pgParams(chrom = "chr4",
                         chromstart = 76173717,
                         chromend = 76253717,
                         assembly = "hg38",
                         range = c(0, scale.SCARB2.max))

# Add the genomic label
plotGenomeLabel(
  chrom = "chr4",
  chromstart = 76173717,
  chromend = 76253717,
  assembly = "hg38",
  x = 0.75,
  y = 8.125,
  length = 4.0,
  default.units = "inches",
  scale = "Mb")

# ATACseq signal smNPC_I
ATAC_smNPC_I_SCARB2 = plotSignal(
  data = bw.SCARB2.smNPC_1, 
  params = region.p.SCARB2,
  fill = "#D3D3D3", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 8.375, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_smNPC_I_SCARB2, 
          at = c(0, scale.SCARB2.max),
          fontsize = 6)

# ATACseq signal smNPC_II
ATAC_smNPC_II_SCARB2 = plotSignal(
  data = bw.SCARB2.smNPC_2, 
  params = region.p.SCARB2,
  fill = "#C0C0C0", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 8.375, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_smNPC_II_SCARB2, 
          at = c(0, scale.SCARB2.max),
          fontsize = 6)

# ATACseq signal smNPC_III
ATAC_smNPC_III_SCARB2 = plotSignal(
  data = bw.SCARB2.smNPC_3, 
  params = region.p.SCARB2,
  fill = "#A9A9A9", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 8.375, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_smNPC_III_SCARB2, 
          at = c(0, scale.SCARB2.max),
          fontsize = 6)

# Add the SNP position that zoom in the TFBS
SNP_param_SCARB2 = pgParams(chrom = "chr4",
                     chromstart = 76213717,
                     chromend = 76213717,
                     assembly = "hg38")

annoHighlight(plot = ATAC_smNPC_I_SCARB2, 
              params = SNP_param_SCARB2,
              fill = "#404788FF",
              y = 8.375, 
              height = 1.875, 
              just = c("left", "top"), 
              default.units = "inches")

annoZoomLines(plot = ATAC_smNPC_I_SCARB2, 
              params = SNP_param_SCARB2,
              y0 = 10.25, 
              x1 = c(2.25, 3.25), 
              y1 = 10.5, 
              default.units = "inches")

# ATACseq signal TH positive neurons D15 I
ATAC_TH.pos_D15_I_SCARB2 = plotSignal(
  data = bw.SCARB2.TH.pos_D15_1, 
  params = region.p.SCARB2,
  fill = "#B22222", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 8.75, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D15_I_SCARB2, 
          at = c(0, scale.SCARB2.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D15 II
ATAC_TH.pos_D15_II_SCARB2 = plotSignal(
  data = bw.SCARB2.TH.pos_D15_2, 
  params = region.p.SCARB2,
  fill = "#A52A2A", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 8.75, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D15_II_SCARB2, 
          at = c(0, scale.SCARB2.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D15 III
ATAC_TH.pos_D15_III_SCARB2 = plotSignal(
  data = bw.SCARB2.TH.pos_D15_3, 
  params = region.p.SCARB2,
  fill = "#8B0000", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 8.75, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D15_III_SCARB2, 
          at = c(0, scale.SCARB2.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D30 I
ATAC_TH.pos_D30_I_SCARB2 = plotSignal(
  data = bw.SCARB2.TH.pos_D30_1, 
  params = region.p.SCARB2,
  fill = "#FF6347", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 9.125, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D30_I_SCARB2, 
          at = c(0, scale.SCARB2.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D30 II
ATAC_TH.pos_D30_II_SCARB2 = plotSignal(
  data = bw.SCARB2.TH.pos_D30_2, 
  params = region.p.SCARB2,
  fill = "#FF0000", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 9.125, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D30_II_SCARB2, 
          at = c(0, scale.SCARB2.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D30 III
ATAC_TH.pos_D30_III_SCARB2 = plotSignal(
  data = bw.SCARB2.TH.pos_D30_3, 
  params = region.p.SCARB2,
  fill = "#DC143C", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 9.125, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D30_III_SCARB2, 
          at = c(0, scale.SCARB2.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D50 I
ATAC_TH.pos_D50_I_SCARB2 = plotSignal(
  data = bw.SCARB2.TH.pos_D50_1, 
  params = region.p.SCARB2,
  fill = "#FFA07A", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 9.5, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D50_I_SCARB2, 
          at = c(0, scale.SCARB2.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D50 II
ATAC_TH.pos_D50_II_SCARB2 = plotSignal(
  data = bw.SCARB2.TH.pos_D50_2, 
  params = region.p.SCARB2,
  fill = "#FA8072", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 9.5, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D50_II_SCARB2, 
          at = c(0, scale.SCARB2.max),
          fontsize = 6)

# ATACseq signal TH positive neurons D50 III
ATAC_TH.pos_D50_III_SCARB2 = plotSignal(
  data = bw.SCARB2.TH.pos_D50_3, 
  params = region.p.SCARB2,
  fill = "#E9967A", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 9.5, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = ATAC_TH.pos_D50_III_SCARB2, 
          at = c(0, scale.SCARB2.max),
          fontsize = 6)

# Gene track
plotGenes(
  chrom = "chr4",
  chromstart = 76173717,
  chromend = 76253717,
  assembly = assembly(Genome = "hg38refGene", 
                      TxDb = "TxDb.Hsapiens.UCSC.hg38.refGene", 
                      OrgDb = "org.Hs.eg.db"),
  fontcolor = "black",
  fill = "black",
  x = 0.75,
  y = 9.875,
  width = 4.0,
  height = 0.375,
  just = c("left", "top"),
  default.units = "inches")

# Add the samples name
plotText(label = "smNPC", 
         fontsize = 6, 
         fontcolor = "#A9A9A9",
         x = 0.75, 
         y = 8.375, 
         just = c("left", "top"),
         default.units = "inches")

plotText(label = "mDAN D15", 
         fontsize = 6, 
         fontcolor = "#B22222",
         x = 0.75, 
         y = 8.75, 
         just = c("left", "top"),
         default.units = "inches")

plotText(label = "mDAN D30", 
         fontsize = 6, 
         fontcolor = "#DC143C",
         x = 0.75, 
         y = 9.125, 
         just = c("left", "top"),
         default.units = "inches")

plotText(label = "mDAN D50", 
         fontsize = 6, 
         fontcolor = "#E9967A",
         x = 0.75, 
         y = 9.5, 
         just = c("left", "top"),
         default.units = "inches")

# text F
plotText(label = "F", 
         fontsize = 12,
         x = 5.0, 
         y = 8.0, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

SCARB2.NR2C2.exp = ggboxplot(dat.RPKM.filt %>% 
                            dplyr::filter(gene_name %in% c("SCARB2", "NR2C2", "FAM47E")) %>% 
                              mutate(gene_name = factor(gene_name,
                                                        levels = c("SCARB2", "NR2C2", "FAM47E"))),
                          x = "Cond", 
                          y = "RPKM", 
                          add = "jitter", 
                          fill = "Cond", 
                          palette = c("#A9A9A9", "#B22222", "#DC143C", "#E9967A"),
                          facet.by = "gene_name",
                          xlab = "", 
                          ylab = "Reads Per Kilobase Million (RPKM)") +
  scale_x_discrete(labels = c("smNPC",
                              "mDAN D15",
                              "mDAN D30",
                              "mDAN D50")) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))

# Place the plot
plotGG(plot = SCARB2.NR2C2.exp, 
       x = 5.25, 
       y = 7.875, 
       width = 3, 
       height = 3.5,
       just = c("left", "top"), 
       default.units = "inches")

# Add the transcription factor binding motif (TFBS) for NR2C2
pfm.NR2C2 = getMatrixByID(JASPAR2020, 
                         ID = "MA1536.1")

# Reverse to position weight matrix as the SNP is on the reverse strand
pfm.NR2C2 = reverseComplement(pfm.NR2C2)

NR2C2 = new.env()

NR2C2$NR2C2 = pfm.NR2C2@profileMatrix

NR2C2 = as.list(NR2C2)
NR2C2_TFBS = ggseqlogo(NR2C2)

# Place the plot
plotGG(plot = NR2C2_TFBS, 
       x = 1.6875, 
       y = 10.375, 
       width = 1.5, 
       height = 1.5,
       just = c("left", "top"), 
       default.units = "inches")

dev.off()
```

### FIGURE 3
```{r fig3, eval = FALSE, echo = TRUE}
########## FIGURE 3 ##########
##############################
# Save as a PDF and as TIF
pdf("/home/vagrant/Manuscript_1/FIGURE3/FIGURE3.pdf",
    paper = "a4",
    family = "Helvetica")

# Save as TIFF, 300 dpi
tiff("/home/vagrant/Manuscript_1/FIGURE3/FIGURE3.tiff",
     width = 8.27,
     height = 11.67,
     units = "in",
     res = 300)

# Create a A4 blank page
pageCreate(width = 8.27, 
           height = 11.67, 
           default.units = "inches",
           showGuides = TRUE)

# text Figure 3
plotText(label = "Figure 3", 
         fontsize = 14,
         fontface = "bold",
         fontfamily = "Helvetica",
         x = 0.25, 
         y = 0.25, 
         just = "left", 
         default.units = "inches")

#### PANEL A - Scheme of transduction for knockdown and luciferase
# text A
plotText(label = "A", 
         fontsize = 12,
         fontface = "bold",
         fontfamily = "Helvetica",
         x = 0.25, 
         y = 0.5, 
         just = "left", 
         default.units = "inches")

#### PANEL B - luciferase barplots with statistics
# text B
plotText(label = "B", 
         fontsize = 12,
         fontface = "bold",
         fontfamily = "Helvetica",
         x = 0.25, 
         y = 2.0, 
         just = "left", 
         default.units = "inches")

Gluc_signal = tibble(Sample = c("BAG3-WT", "BAG3-MUT", "IDUA-WT", "IDUA-MUT", "SCARB2-WT", "SCARB2-MUT", "miniCMV", "Neg_CTRL"),
                     GLUC_n1_rep1 = c(319, 240, 154, 196, 379, 211, 431, 33),
                     GLUC_n1_rep2 = c(323, 237, 161, 205, 371, 224, 459, 33),
                     GLUC_n2_rep1 = c(201, 143, 121, 128, 278, 172, 253, 31),
                     GLUC_n2_rep2 = c(191, 149, 119, 126, 278, 162, 249, 31),
                     GLUC_n3_rep1 = c(249, 146, 136, 142, 223, 175, 212, 32),
                     GLUC_n3_rep2 = c(203, 152, 121, 125, 209, 168, 190, 33),
                     GLUC_n4_rep1 = c(348, 333, 220, 185, 396, 244, 319, 31),
                     GLUC_n4_rep2 = c(322, 306, 229, 186, 369, 227, 294, 32),
                     SEAP_n1_rep1 = c(72, 78, 54, 67, 99, 75, 128, 32),
                     SEAP_n1_rep2 = c(69, 70, 49, 66, 97, 79, 132, 29),
                     SEAP_n2_rep1 = c(50, 52, 45, 45, 76, 64, 80, 32),
                     SEAP_n2_rep2 = c(50, 52, 45, 51, 85, 63, 80, 32),
                     SEAP_n3_rep1 = c(52, 50, 43, 50, 66, 55, 68, 32),
                     SEAP_n3_rep2 = c(49, 53, 48, 46, 66, 62, 69, 31),
                     SEAP_n4_rep1 = c(64, 75, 58, 54, 103, 75, 111, 33),
                     SEAP_n4_rep2 = c(61, 74, 62, 56, 108, 77, 111, 31)) %>% 
  pivot_longer(!Sample, names_to = "Meas", values_to = "Abs_val") %>% 
  mutate(Meas2 = gsub("^([^_]*_[^_]*)_.*", "\\1", Meas)) %>% 
  group_by(Sample, Meas2) %>% 
  mutate(AVG = mean(Abs_val, na.rm = TRUE))

# Calculate ratio Gluc over SEAP
Gluc_signal_ratio = Gluc_signal %>% 
  dplyr::select(-Meas, -Abs_val) %>% 
  distinct(Meas2, .keep_all = TRUE) %>% 
  pivot_wider(names_from = "Meas2", values_from = "AVG") %>% 
  mutate(Ratio_within_bio_rep1 = GLUC_n1/SEAP_n1,
         Ratio_within_bio_rep2 = GLUC_n2/SEAP_n2,
         Ratio_within_bio_rep3 = GLUC_n3/SEAP_n3,
         Ratio_within_bio_rep4 = GLUC_n4/SEAP_n4) %>% 
  dplyr::select(Sample, Ratio_within_bio_rep1:Ratio_within_bio_rep4) %>% 
  pivot_longer(cols = c("Ratio_within_bio_rep1", "Ratio_within_bio_rep2", "Ratio_within_bio_rep3",
                        "Ratio_within_bio_rep4"),
               names_to = "Ratio",
               values_to = "value") %>% 
  mutate(Genes = gsub("(.*)-.*", "\\1", Sample))

# Calculate statistics - T-test
GLUC_stat = compare_means(value ~ Sample,
                          data = Gluc_signal_ratio %>% 
                            ungroup(),
                          method = "t.test")

# Plot for BAG3
GLUC_BAG3 = ggbarplot(Gluc_signal_ratio %>% 
                        dplyr::filter(Sample %in% c("BAG3-WT",
                                                     "BAG3-MUT",
                                                     "miniCMV",
                                                     "Neg_CTRL")), 
                      x = "Sample", 
                      y = "value",
                      width = 0.25,
                      add = c("mean_se", 
                              "jitter"),
                      fill = "grey",
                      xlab = "",
                      ylab = "Ratio Gluc/SEAP",
                      position = position_dodge(0.9)) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1.0, size = 9, face = "bold", family = "Helvetica"),
        axis.text.y = element_text(size = 9, face = "bold"),
        axis.title.y = element_text(size = 10, face = "bold")) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 6.0)) +
  scale_x_discrete(labels = c("BAG3 other allele",
                              "BAG3 effect allele",
                              "Positive control",
                              "Negative control")) +
  stat_pvalue_manual(GLUC_stat %>% 
                       dplyr::filter(group1 == "BAG3-WT",
                              group2 == "BAG3-MUT"),
                     label = "p = {p.format}",
                     y.position = 5.6,
                     size = 3.0)

# Place the plot
plotGG(plot = GLUC_BAG3, 
       x = 0.50, 
       y = 2.25, 
       width = 2, 
       height = 4.25,
       just = c("left", "top"), 
       default.units = "inches")

# Plot for IDUA
GLUC_IDUA = ggbarplot(Gluc_signal_ratio %>% 
                        dplyr::filter(Sample %in% c("IDUA-WT",
                                                    "IDUA-MUT",
                                                    "miniCMV",
                                                    "Neg_CTRL")), 
                      x = "Sample", 
                      y = "value",
          width = 0.25,
          add = c("mean_se", 
                  "jitter"),
          fill = "grey",
          xlab = "",
          ylab = "Ratio Gluc/SEAP",
          position = position_dodge(0.9)) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1.0, size = 9, face = "bold"),
        axis.text.y = element_text(size = 9, face = "bold"),
        axis.title.y = element_text(size = 10, face = "bold")) +
  scale_x_discrete(labels = c("IDUA other allele",
                              "IDUA effect allele",
                              "Positive control",
                              "Negative control")) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 6.0)) +
  stat_pvalue_manual(GLUC_stat %>% 
                       dplyr::filter(group1 == "IDUA-WT",
                              group2 == "IDUA-MUT"),
                     label = "p = {p.format}",
                     y.position = 4.0,
                     size = 3.0)

# Place the plot
plotGG(plot = GLUC_IDUA, 
       x = 3.0, 
       y = 2.25, 
       width = 2, 
       height = 4.25,
       just = c("left", "top"), 
       default.units = "inches")

# Plot for SCARB2
GLUC_SCARB2 = ggbarplot(Gluc_signal_ratio %>% 
                          dplyr::filter(Sample %in% c("SCARB2-WT",
                                                     "SCARB2-MUT",
                                                     "miniCMV",
                                                     "Neg_CTRL")), 
                        x = "Sample", 
                        y = "value",
          width = 0.25,
          add = c("mean_se", 
                  "jitter"),
          fill = "grey",
          xlab = "",
          ylab = "Ratio Gluc/SEAP",
          position = position_dodge(0.9)) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1.0, size = 9, face = "bold"),
        axis.text.y = element_text(size = 9, face = "bold"),
        axis.title.y = element_text(size = 10, face = "bold")) +
  scale_x_discrete(labels = c("SCARB2 other allele",
                              "SCARB2 effect allele",
                              "Positive control",
                              "Negative control")) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 6.0)) +
  stat_pvalue_manual(GLUC_stat %>% 
                       dplyr::filter(group1 == "SCARB2-WT",
                              group2 == "SCARB2-MUT"),
                     label = "p = {p.format}",
                     y.position = 4.0,
                      size = 3.0)

# Place the plot
plotGG(plot = GLUC_SCARB2, 
       x = 5.5, 
       y = 2.25, 
       width = 2, 
       height = 4.25,
       just = c("left", "top"), 
       default.units = "inches")

#### PANEL C - knockdown barplots with statistics
# text C
plotText(label = "C", 
         fontsize = 12,
         x = 0.25, 
         y = 6.75, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# Knockdown of LHX1 and quantification of LHX1 and BAG3 expression 3 days post transduction
KD.BAG3.LHX1.3days = tibble(Sple = rep("shLHX1", 6),
                            Gene = c(rep("LHX1", 3),
                                     rep("BAG3", 3)),
                            Exp = c(0.411253539,
                                    0.359459217,
                                    0.367190337,
                                    0.972688658,
                                    0.965902853,
                                    0.88813445))

# Stats
KD.BAG3.LHX1.3days.stats = tibble(Sple = c(rep("shLHX1", 6),
                                           rep("shCTRL", 6)),
                                  Gene = rep(c(rep("LHX1", 3),
                                           rep("BAG3", 3)), 2),
                                  Exp = c(9.5167,
                                          8.39675,
                                          8.1855,
                                          9.13605,
                                          8.33795,
                                          8.76415,
                                          8.2348,
                                          6.92065,
                                          6.7401,
                                          9.0961,
                                          8.2879,
                                          8.593))

stat_LHX1.KD = compare_means(Exp ~ Sple,
                          data = KD.BAG3.LHX1.3days.stats,
                          group.by = "Gene",
                          method = "t.test",
                          paired = TRUE)
# Plot
pl.KD_BAG3_LHX1 = KD.BAG3.LHX1.3days %>% 
  mutate(Gene = factor(Gene, levels = c("LHX1", "BAG3"))) %>% 
  ggbarplot(.,
            x = "Gene",
            y = "Exp",
            fill = c("Gene"),
            position = position_dodge(0.9),
            palette = c("#440154FF", "#287C8EFF"),
            add = c("mean_se", "jitter"),
            xlab = "",
            ylab = "Relative expression to shSCRAMBLE",
            title = "shLHX1 - 3 days") +
  geom_hline(yintercept = 1.0, lty = "dashed", color = "black") +
  theme(axis.title.y = element_text(face = "bold", size = 10),
        axis.text.x = element_text(face = "bold.italic", size = 10, angle = 90, family = "Helvetica"),
        axis.text.y = element_text(face = "bold", size = 10),
        legend.position = "none",
        plot.title = element_text(face = "bold", size = 10, hjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 1.8)) +
  stat_pvalue_manual(stat_LHX1.KD,
                     label = "p = {p.format}",
                     x = "Gene",
                     y.position = 0.5,
                     size = 3.0,
                     hide.ns = TRUE)

plotGG(plot = pl.KD_BAG3_LHX1, 
       x = 0.5, 
       y = 7.0, 
       width = 1.75, 
       height = 4.25,
       just = c("left", "top"), 
       default.units = "inches")

# Knockdown of ZBTB14 and quantification of ZBTB14 and IDUA expression 3 days post transduction
KD.IDUA.ZBTB14.3days = tibble(Sple = rep("shZBTB14", 6),
                            Gene = c(rep("ZBTB14", 3),
                                     rep("IDUA", 3)),
                            Exp = c(0.334134302,
                                    0.316834205,
                                    0.294206291,
                                    0.663813558,
                                    1.259979272,
                                    1.774054093))

# Stats
KD.IDUA.ZBTB14.3days.stats = tibble(Sple = c(rep("shZBTB14", 6),
                                           rep("shCTRL", 6)),
                                  Gene = rep(c(rep("ZBTB14", 3),
                                           rep("IDUA", 3)), 2),
                                  Exp = c(11.39125,
                                          10.4813,
                                          10.5062,
                                          11.56275,
                                          10.1455,
                                          9.76485,
                                          9.80975,
                                          8.8231,
                                          8.7411,
                                          10.9716,
                                          10.4789,
                                          10.5919))

stat_ZBTB14.KD = compare_means(Exp ~ Sple,
                          data = KD.IDUA.ZBTB14.3days.stats,
                          group.by = "Gene",
                          method = "t.test",
                          paired = TRUE)

# Plot
pl.KD_IDUA_ZBTB14 = KD.IDUA.ZBTB14.3days %>% 
  mutate(Gene = factor(Gene, levels = c("ZBTB14", "IDUA"))) %>% 
  ggbarplot(.,
            x = "Gene",
            y = "Exp",
            fill = c("Gene"),
            position = position_dodge(0.9),
            palette = c("#440154FF", "#287C8EFF"),
            add = c("mean_se", "jitter"),
            xlab = "",
            ylab = "Relative expression to shSCRAMBLE",
            title = "shZBTB14 - 3 days") +
  geom_hline(yintercept = 1.0, lty = "dashed", color = "black") +
  theme(axis.title.y = element_text(face = "bold", size = 10),
        axis.text.x = element_text(face = "bold.italic", size = 10, angle = 90, family = "Helvetica"),
        axis.text.y = element_text(face = "bold", size = 10),
        legend.position = "none",
        plot.title = element_text(face = "bold", size = 10, hjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 1.8)) +
  stat_pvalue_manual(stat_ZBTB14.KD,
                     label = "p = {p.format}",
                     x = "Gene",
                     y.position = 0.5,
                     size = 3.0,
                     hide.ns = TRUE)

plotGG(plot = pl.KD_IDUA_ZBTB14, 
       x = 2.375, 
       y = 7.0, 
       width = 1.75, 
       height = 4.25,
       just = c("left", "top"), 
       default.units = "inches")

# Knockdown of NR2C2 and quantification of NR2C2 and SCARB2 expression 3 days post transduction
KD.SCARB2.NR2C2.3days = tibble(Sple = rep("shNR2C2", 6),
                            Gene = c(rep("NR2C2", 3),
                                     rep("SCARB2", 3)),
                            Exp = c(0.345210918,
                                    0.512864636,
                                    0.548665969,
                                    1.239922544,
                                    1.130295504,
                                    1.170885498))

# Stats
KD.SCARB2.NR2C2.3days.stats = tibble(Sple = c(rep("shNR2C2", 6),
                                           rep("shCTRL", 6)),
                                  Gene = rep(c(rep("NR2C2", 3),
                                           rep("SCARB2", 3)), 2),
                                  Exp = c(9.43575,
                                          9.3043,
                                          9.263,
                                          5.6841,
                                          5.92295,
                                          5.9289,
                                          7.9013,
                                          8.34095,
                                          8.397,
                                          5.99435,
                                          6.09965,
                                          6.1565))

stat_NR2C2.KD = compare_means(Exp ~ Sple,
                          data = KD.SCARB2.NR2C2.3days.stats,
                          group.by = "Gene",
                          method = "t.test",
                          paired = TRUE)

# Plot
pl.KD_SCARB2_NR2C2 = KD.SCARB2.NR2C2.3days %>% 
  mutate(Gene = factor(Gene, levels = c("NR2C2", "SCARB2"))) %>% 
  ggbarplot(.,
            x = "Gene",
            y = "Exp",
            fill = c("Gene"),
            position = position_dodge(0.9),
            palette = c("#440154FF", "#287C8EFF"),
            add = c("mean_se", "jitter"),
            xlab = "",
            ylab = "Relative expression to shSCRAMBLE",
            title = "shNR2C2 - 3 days") +
  geom_hline(yintercept = 1.0, lty = "dashed", color = "black") +
  theme(axis.title.y = element_text(face = "bold", size = 10),
        axis.text.x = element_text(face = "bold.italic", size = 10, angle = 90, family = "Helvetica"),
        axis.text.y = element_text(face = "bold", size = 10),
        legend.position = "none",
        plot.title = element_text(face = "bold", size = 10, hjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 1.8)) +
  stat_pvalue_manual(stat_NR2C2.KD,
                     label = "p = {p.format}",
                     x = "Gene",
                     y.position = 1.5,
                     size = 3.0)

plotGG(plot = pl.KD_SCARB2_NR2C2, 
       x = 4.25, 
       y = 7.0, 
       width = 1.75, 
       height = 4.25,
       just = c("left", "top"), 
       default.units = "inches")

# Knockdown of NR2C2 and quantification of NR2C2 and SCARB2 expression 6 days post transduction
KD.SCARB2.NR2C2.6days = tibble(Sple = rep("shNR2C2", 6),
                            Gene = c(rep("NR2C2", 3),
                                     rep("SCARB2", 3)),
                            Exp = c(0.662710186,
                                    0.456029798,
                                    0.787662775,
                                    1.410640142,
                                    1.425976296,
                                    1.332143962))

# Stats
KD.SCARB2.NR2C2.6days.stats = tibble(Sple = c(rep("shNR2C2", 6),
                                           rep("shCTRL", 6)),
                                  Gene = rep(c(rep("NR2C2", 3),
                                           rep("SCARB2", 3)), 2),
                                  Exp = c(8.34355,
                                          9.0917,
                                          8.32445,
                                          5.3595,
                                          5.19535,
                                          6.5266,
                                          7.75,
                                          7.9589,
                                          7.9801,
                                          5.85585,
                                          5.7073,
                                          6.94035))

stat_NR2C2.KD.6days = compare_means(Exp ~ Sple,
                          data = KD.SCARB2.NR2C2.6days.stats,
                          group.by = "Gene",
                          method = "t.test",
                          paired = TRUE)

# Plot
pl.KD_SCARB2_NR2C2_6days = KD.SCARB2.NR2C2.6days %>% 
  mutate(Gene = factor(Gene, levels = c("NR2C2", "SCARB2"))) %>% 
  ggbarplot(.,
            x = "Gene",
            y = "Exp",
            fill = c("Gene"),
            position = position_dodge(0.9),
            palette = c("#440154FF", "#287C8EFF"),
            add = c("mean_se", "jitter"),
            xlab = "",
            ylab = "Relative expression to shSCRAMBLE",
            title = "shNR2C2 - 6 days") +
  geom_hline(yintercept = 1.0, lty = "dashed", color = "black") +
  theme(axis.title.y = element_text(face = "bold", size = 10),
        axis.text.x = element_text(face = "bold.italic", size = 10, angle = 90, family = "Helvetica"),
        axis.text.y = element_text(face = "bold", size = 10),
        legend.position = "none",
        plot.title = element_text(face = "bold", size = 10, hjust = 0.5)) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 1.8)) +
  stat_pvalue_manual(stat_NR2C2.KD.6days,
                     label = "p = {p.format}",
                     x = "Gene",
                     y.position = 1.5,
                     size = 3.0,
                     hide.ns = TRUE)

plotGG(plot = pl.KD_SCARB2_NR2C2_6days, 
       x = 6.125, 
       y = 7.0, 
       width = 1.75, 
       height = 4.25,
       just = c("left", "top"), 
       default.units = "inches")

dev.off()
```

### FIGURE 4
```{r fig4, eval = FALSE, echo = TRUE}
########## FIGURE 4 ##########
##############################
# Save as a PDF
pdf("/home/vagrant/Manuscript_1/FIGURE4/FIGURE4.pdf",
    width = 8.3, 
    height = 11.7)

# Save as TIFF, 300 dpi
tiff("/home/vagrant/Manuscript_1/FIGURE4/FIGURE4.tiff",
     width = 8.27,
     height = 11.67,
     units = "in",
     res = 300)

# Create a A4 blank page
pageCreate(width = 8.27, 
           height = 11.67, 
           default.units = "inches",
           showGuides = TRUE)

plotText(label = "Figure 4", 
         fontsize = 14,
         fontface = "bold",
         fontfamily = "Helvetica",
         x = 0.25, 
         y = 0.25, 
         just = "left", 
         default.units = "inches")

#### PANEL A - Chromatin imbalance in TH REP1 mCHERRY cell line at rs1465922 
# text A
plotText(label = "A", 
         fontsize = 12,
         fontface = "bold",
         fontfamily = "Helvetica",
         x = 0.25, 
         y = 0.5, 
         just = "left", 
         default.units = "inches")

#FIG 4A
AI_RNA = tibble(Samples = 
                  c("smNPCs", "Neurons positively sorted D15", "Neurons positively sorted D30",
                    "Neurons positively sorted D50", "Astrocytes"),
                A = c(5, 23, 77, 56, 30),
                G = c(8, 4, 62, 34, 47)) %>% 
  gather(-Samples, key = "Base", value = "Reads")

AI_ATAC = tibble(Samples = 
                  c("smNPCs", "mDAN D15", "mDAN D30",
                    "mDAN D50", "Astrocytes"),
                A = c(31, 195, 202, 118, 155),
                G = c(39, 263, 264, 196, 179)) %>% 
  gather(-Samples, key = "Base", value = "Reads")

AI = AI_ATAC %>% 
  bind_rows(AI_RNA) %>% 
  mutate(Cond = rep(c("ATAC", "RNA"), each = 10))

# Good looking plot
chrom_AI = ggbarplot(AI_ATAC, x = "Samples", y = "Reads",
          fill = "Base",
          position = position_dodge(0.7),
          xlab = "", ylab = "Reads",
          #title = "Allelic imbalance at chr4:76213717 - SNP in SCARB2 promoter",
          label = c(0.3, 0.001, 0.004, 1e-5, 0.2,
                    rep("", 5)),
          label.pos = "out",
          lab.size = 3) +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust = 1.0, face = "bold", size = 10),
        axis.text.y = element_text(face = "bold", size = 10),
        axis.title.y = element_text(face = "bold", size = 10),
        legend.text = element_text(face = "bold", size = 10),
        legend.title = element_text(face = "bold", size = 10)) +
        # title = element_text(hjust = 0.5)) +
  scale_fill_manual(values = c("#80A44F", "#427F96"))

plotGG(plot = chrom_AI, 
       x = 0.5, 
       y = 0.75, 
       width = 6, 
       height = 4,
       just = c("left", "top"), 
       default.units = "inches")

# text B
plotText(label = "B", 
         fontsize = 12,
         x = 0.25, 
         y = 5.0, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

GTEX_SNPs_interest = read_delim("/home/vagrant/epifunc/eQTpLot/20230117_GTEX_data.csv",
               delim = ";",
               col_names = TRUE)

GTEX_SNPs_SCARB2_brain = GTEX_SNPs_interest %>% 
  mutate(GTEX_sig = if_else(P.Value < 0.05, "*", "")) %>% 
  filter(str_detect(Tissue, "Brain"),
         Gene.Symbol == "SCARB2")

GTEX_eqtl = ggplot(data = GTEX_SNPs_SCARB2_brain, mapping = aes(x = SNP.Id,
                                                y = Tissue,
                                                fill = NES)) + 
  geom_tile() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, face = "bold", size = 10),
        axis.title.x = element_blank(),
        axis.text.y = element_text(face = "bold", size = 10),
        axis.title.y = element_blank(),
        legend.text = element_text(face = "bold", size = 10),
        legend.title = element_text(face = "bold", size = 10)) +
  scale_fill_gradient2(
    low = muted("blue"),
    mid = "white",
    high = muted("red"),
    midpoint = 0,
    name = "eQTL \neffect size") +
  geom_text(aes(label = GTEX_sig))

plotGG(plot = GTEX_eqtl, 
       x = 0.5, 
       y = 5.25, 
       width = 7, 
       height = 6,
       just = c("left", "top"), 
       default.units = "inches")

dev.off()
```

### FIGURE 5
```{r fig5, eval = FALSE, echo = TRUE}
########## FIGURE 5 ##########
##############################
# Save as a PDF
pdf("/home/vagrant/Manuscript_1/FIGURE5/FIGURE5.pdf",
    width = 8.3, 
    height = 11.7)

# Create a A4 blank page
pageCreate(width = 8.3, 
           height = 11.7, 
           default.units = "inches",
           showGuides = FALSE)

#### PANEL A - 
# text Figure 5
plotText(label = "Figure 5", 
         fontsize = 14,
         x = 0.25, 
         y = 0.25, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")
# text A
plotText(label = "A", 
         fontsize = 12,
         x = 0.25, 
         y = 0.5, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

KD_TH_lev = tibble(Sple = rep("shNR2C2", 3),
                   Rep = 1:3,
                   days = "6 days",
                   Gene = "TH",
                   Exp = c(1.705920009,
                           2.196566231,
                           1.107610012))

#TH levels in knockdown of NR2C2 6 days#
TH_lev_KD = KD_TH_lev %>% 
  ggbarplot(.,
            x = "Gene",
            y = "Exp",
            fill = c("days"),
            facet.by = "Sple",
            position = position_dodge(0.9),
            palette = c("#287C8EFF"),
            add = c("mean_se"),
            xlab = "",
            ylab = "Relative expression to shSCRAMBLE") +
  geom_hline(yintercept = 1.0, lty = "dashed", color = "black") +
  theme(axis.title.y = element_text(face = "bold", size = 10),
        axis.text.x = element_text(face = "bold", size = 10, angle = 90),
        axis.text.y = element_text(face = "bold", size = 10),
        legend.position = "none")

plotGG(plot = TH_lev_KD, 
       x = 0.5, 
       y = 0.5, 
       width = 2, 
       height = 4.25,
       just = c("left", "top"), 
       default.units = "inches")

### DEXOMAG - 3 biological repliactes ###
GCase = read_excel("/Volumes/deborah.gerard/Documents/Wetlab/epifunc/DEXOMAG_shSCARB2_shNR2C2_shCTRL_N1_N2_N3/20231214_DEXOMAG_KD_SCARB2_NR2C2_SCRAMBLE_N1_N2_N3_ANALYSIS.xlsx")

GCase %>% 
  dplyr::select(GCase, `GCase MEAN`, ...10, ...11) %>% 
  drop_na() %>% 
  dplyr::rename(Sample = GCase,
                N1 = `GCase MEAN`,
                N2 = ...10,
                N3 = ...11) %>% 
  pivot_longer(N1:N3, names_to = "Rep.bio", values_to = "norm.fluo.value") %>% 
  mutate(norm.fluo.value = as.numeric(norm.fluo.value)) %>% 
  #dplyr::filter(Sample != "Whole Gcase NT") %>% 
  ggbarplot(.,
            x = "Sample",
            y = "norm.fluo.value",
            add = c("mean_se", "jitter"),
            xlab = "",
            ylab = "GCase activity (normalised fluorescence values)") +
  scale_x_discrete(labels = c("SCARB2 KD",
                              "NR2C2 KD",
                              "SCRAMBLE",
                              "Not transduced + dexomag",
                              "Not transduced - dexomag",
                              "Whole cell lysate")) +
  theme(axis.text.x = element_text(angle = 45, face = "bold", hjust = 1.0),
        axis.text.y = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold")) +
  stat_compare_means(comparisons = list(c("shSCARB2", "shSCRAMBLE")),
                     method = "t.test",
                     size = 4,
                     label = "p.format") +
  stat_compare_means(comparisons = list(c("shNR2C2", "shSCRAMBLE")),
                     method = "t.test",
                     size = 4,
                     label = "p.format")

dev.off()
```

### SUPPLEMENTARY FIGURE 1
IHEC consortium has one sample of substantia nigra for H3K27me3, H3K9me3, H3K4me1, H3K4me3 and H3K36me3 from a healthy brain for BAG3, IDUA and SCARB2 loci
```{r suppFig1, eval = FALSE, echo = TRUE}
############################################
########## SUPPLEMENTARY FIGURE 1 ##########
############################################

# Save as a PDF
pdf("/home/vagrant/Manuscript_1/SUPPLEMENTARY_FIGURE1/SUPPLEMENTARY_FIGURE1.pdf",
    width = 8.3, 
    height = 11.7)

# Create a A4 blank page
pageCreate(width = 8.3, 
           height = 11.7, 
           default.units = "inches",
           showGuides = TRUE)

#### PANEL A - 
# text Supplementary Figure 1
plotText(label = "Supplementary Figure 1", 
         fontsize = 14,
         x = 0.25, 
         y = 0.25, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")
# text A
plotText(label = "A", 
         fontsize = 12,
         x = 0.25, 
         y = 0.5, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# ChIP-seq signal BAG3
# path where bigwig files are
bw.IHEC.path = "/home/vagrant/Documents/IHEC/ChIP_Seq/BIGWIG/"

# Reads BIGWIG file at BAG3 location - H3K4me1
bw.BAG3.H3K4me1 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.6a1889bf-7312-4e41-b1cc-e2f3be7a8076.fc.signal.bigwig"),
                               chrom = "chr10",
                               chromstart = 119611305,
                               chromend = 119691505)

# Reads BIGWIG file at BAG3 location - H3K4me3
bw.BAG3.H3K4me3 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.4cb9300f-9097-4374-8f2a-4afd8ff855fb.fc.signal.bigwig"),
                               chrom = "chr10",
                               chromstart = 119611305,
                               chromend = 119691505)

# Reads BIGWIG file at BAG3 location - H3K36me3
bw.BAG3.H3K36me3 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.6586e624-852f-4cac-83df-2325cee19699.fc.signal.bigwig"),
                               chrom = "chr10",
                               chromstart = 119611305,
                               chromend = 119691505)

# Reads BIGWIG file at BAG3 location - H3K9me3
bw.BAG3.H3K9me3 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.e8fe7664-906e-4d87-8f3a-6ce88dbeadc6.fc.signal.bigwig"),
                               chrom = "chr10",
                               chromstart = 119611305,
                               chromend = 119691505)

# Reads BIGWIG file at BAG3 location - H3K27me3
bw.BAG3.H3K27me3 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.5f5024a9-2e3f-4527-82ec-0ed04bf4a848.fc.signal.bigwig"),
                               chrom = "chr10",
                               chromstart = 119611305,
                               chromend = 119691505)

# Add a scale next to the bigwig files - check the maximum to choose
scale.IHEC.BAG3.max = max(c(bw.BAG3.H3K4me1$score,
                              bw.BAG3.H3K4me3$score,
                              bw.BAG3.H3K36me3$score,
                              bw.BAG3.H3K9me3$score,
                              bw.BAG3.H3K27me3$score)) %>% 
  round(.,
        digits = 1.0)

print(scale.IHEC.BAG3.max)

# Define parameters for the regions
region.p.IHEC.BAG3 = pgParams(chrom = "chr10",
                         chromstart = 119611405,
                         chromend = 119691405,
                         assembly = "hg38",
                         range = c(0, scale.IHEC.BAG3.max))

# Add the genomic label
plotGenomeLabel(
  chrom = "chr10",
  chromstart = 119611405,
  chromend = 119691405,
  assembly = "hg38",
  x = 0.75,
  y = 0.625,
  length = 4.0,
  default.units = "inches",
  scale = "Mb")

# ChIPseq signal H3K4me1
H3K4me1_BAG3 = plotSignal(
  data = bw.BAG3.H3K4me1, 
  params = region.p.IHEC.BAG3,
  fill = "#b5de2b", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 0.875, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K4me1_BAG3, 
          at = c(0, scale.IHEC.BAG3.max),
          fontsize = 6)

# ChIPseq signal H3K4me3
H3K4me3_BAG3 = plotSignal(
  data = bw.BAG3.H3K4me3, 
  params = region.p.IHEC.BAG3,
  fill = "#35b779", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 1.25, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K4me3_BAG3, 
          at = c(0, scale.IHEC.BAG3.max),
          fontsize = 6)

# ChIPseq signal H3K36me3
H3K36me3_BAG3 = plotSignal(
  data = bw.BAG3.H3K36me3, 
  params = region.p.IHEC.BAG3,
  fill = "#26828e", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 1.625, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K36me3_BAG3, 
          at = c(0, scale.IHEC.BAG3.max),
          fontsize = 6)

# ChIPseq signal H3K9me3
H3K9me3_BAG3 = plotSignal(
  data = bw.BAG3.H3K9me3, 
  params = region.p.IHEC.BAG3,
  fill = "#3e4989", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 2.0, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K9me3_BAG3, 
          at = c(0, scale.IHEC.BAG3.max),
          fontsize = 6)

# ChIPseq signal H3K27me3
H3K27me3_BAG3 = plotSignal(
  data = bw.BAG3.H3K27me3, 
  params = region.p.IHEC.BAG3,
  fill = "#440154", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 2.375, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K27me3_BAG3, 
          at = c(0, scale.IHEC.BAG3.max),
          fontsize = 6)

# Gene track
plotGenes(
  chrom = "chr10",
  chromstart = 119611405,
  chromend = 119691405,
  assembly = assembly(Genome = "hg38refGene", 
                      TxDb = "TxDb.Hsapiens.UCSC.hg38.refGene", 
                      OrgDb = "org.Hs.eg.db"),
  fontcolor = "black",
  fill = "black",
  x = 0.75,
  y = 2.75,
  width = 4.0,
  height = 0.375,
  just = c("left", "top"),
  default.units = "inches")

# Add the samples name
plotText(label = "H3K4me1", 
         fontsize = 6, 
         fontcolor = "#b5de2b",
         x = 0.75,
         y = 0.875, 
         just = c("left", "top"),
         default.units = "inches")

# Add the samples name
plotText(label = "H3K4me3", 
         fontsize = 6, 
         fontcolor = "#35b779",
         x = 0.75,
         y = 1.25, 
         just = c("left", "top"),
         default.units = "inches")

# Add the samples name
plotText(label = "H3K36me3", 
         fontsize = 6, 
         fontcolor = "#26828e",
         x = 0.75,
         y = 1.625, 
         just = c("left", "top"),
         default.units = "inches")

# Add the samples name
plotText(label = "H3K9me3", 
         fontsize = 6, 
         fontcolor = "#3e4989",
         x = 0.75,
         y = 2.0, 
         just = c("left", "top"),
         default.units = "inches")

# Add the samples name
plotText(label = "H3K27me3", 
         fontsize = 6, 
         fontcolor = "#440154",
         x = 0.75,
         y = 2.375, 
         just = c("left", "top"),
         default.units = "inches")

# Reads BIGWIG file at IDUA location - H3K4me1
bw.IDUA.H3K4me1 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.6a1889bf-7312-4e41-b1cc-e2f3be7a8076.fc.signal.bigwig"),
                               chrom = "chr4",
                               chromstart = 947008,
                               chromend = 1027208)

# Reads BIGWIG file at IDUA location - H3K4me3
bw.IDUA.H3K4me3 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.4cb9300f-9097-4374-8f2a-4afd8ff855fb.fc.signal.bigwig"),
                               chrom = "chr4",
                               chromstart = 947008,
                               chromend = 1027208)

# Reads BIGWIG file at IDUA location - H3K36me3
bw.IDUA.H3K36me3 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.6586e624-852f-4cac-83df-2325cee19699.fc.signal.bigwig"),
                               chrom = "chr4",
                               chromstart = 947008,
                               chromend = 1027208)

# Reads BIGWIG file at IDUA location - H3K9me3
bw.IDUA.H3K9me3 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.e8fe7664-906e-4d87-8f3a-6ce88dbeadc6.fc.signal.bigwig"),
                               chrom = "chr4",
                               chromstart = 947008,
                               chromend = 1027208)

# Reads BIGWIG file at IDUA location - H3K27me3
bw.IDUA.H3K27me3 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.5f5024a9-2e3f-4527-82ec-0ed04bf4a848.fc.signal.bigwig"),
                               chrom = "chr4",
                               chromstart = 947008,
                               chromend = 1027208)

# Add a scale next to the bigwig files - check the maximum to choose
scale.IHEC.IDUA.max = max(c(bw.IDUA.H3K4me1$score,
                              bw.IDUA.H3K4me3$score,
                              bw.IDUA.H3K36me3$score,
                              bw.IDUA.H3K9me3$score,
                              bw.IDUA.H3K27me3$score)) %>% 
  round(.,
        digits = 1.0)

print(scale.IHEC.IDUA.max)

# Define parameters for the regions
region.p.IHEC.IDUA = pgParams(chrom = "chr4",
                         chromstart = 947108,
                         chromend = 1027108,
                         assembly = "hg38",
                         range = c(0, scale.IHEC.IDUA.max))

# Add the genomic label
plotGenomeLabel(
  chrom = "chr4",
  chromstart = 947108,
  chromend = 1027108,
  assembly = "hg38",
  x = 0.75,
  y = 3.5,
  length = 4.0,
  default.units = "inches",
  scale = "Mb")

# ChIPseq signal H3K4me1
H3K4me1_IDUA = plotSignal(
  data = bw.IDUA.H3K4me1, 
  params = region.p.IHEC.IDUA,
  fill = "#b5de2b", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 3.75, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K4me1_IDUA, 
          at = c(0, scale.IHEC.IDUA.max),
          fontsize = 6)

# ChIPseq signal H3K4me3
H3K4me3_IDUA = plotSignal(
  data = bw.IDUA.H3K4me3, 
  params = region.p.IHEC.IDUA,
  fill = "#35b779", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 4.125, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K4me3_IDUA, 
          at = c(0, scale.IHEC.IDUA.max),
          fontsize = 6)

# ChIPseq signal H3K36me3
H3K36me3_IDUA = plotSignal(
  data = bw.IDUA.H3K36me3, 
  params = region.p.IHEC.IDUA,
  fill = "#26828e", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 4.5, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K36me3_IDUA, 
          at = c(0, scale.IHEC.IDUA.max),
          fontsize = 6)

# ChIPseq signal H3K9me3
H3K9me3_IDUA = plotSignal(
  data = bw.IDUA.H3K9me3, 
  params = region.p.IHEC.IDUA,
  fill = "#3e4989", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 4.875, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K9me3_IDUA, 
          at = c(0, scale.IHEC.IDUA.max),
          fontsize = 6)

# ChIPseq signal H3K27me3
H3K27me3_IDUA = plotSignal(
  data = bw.IDUA.H3K27me3, 
  params = region.p.IHEC.IDUA,
  fill = "#440154", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 5.25, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K27me3_IDUA, 
          at = c(0, scale.IHEC.IDUA.max),
          fontsize = 6)

# Gene track
plotGenes(
  chrom = "chr4",
  chromstart = 947108,
  chromend = 1027108,
  assembly = assembly(Genome = "hg38refGene", 
                      TxDb = "TxDb.Hsapiens.UCSC.hg38.refGene", 
                      OrgDb = "org.Hs.eg.db"),
  fontcolor = "black",
  fill = "black",
  x = 0.75,
  y = 5.625,
  width = 4.0,
  height = 0.375,
  just = c("left", "top"),
  default.units = "inches")

# Add the samples name
plotText(label = "H3K4me1", 
         fontsize = 6, 
         fontcolor = "#b5de2b",
         x = 0.75,
         y = 3.75, 
         just = c("left", "top"),
         default.units = "inches")

# Add the samples name
plotText(label = "H3K4me3", 
         fontsize = 6, 
         fontcolor = "#35b779",
         x = 0.75,
         y = 4.125, 
         just = c("left", "top"),
         default.units = "inches")

# Add the samples name
plotText(label = "H3K36me3", 
         fontsize = 6, 
         fontcolor = "#26828e",
         x = 0.75,
         y = 4.5, 
         just = c("left", "top"),
         default.units = "inches")

# Add the samples name
plotText(label = "H3K9me3", 
         fontsize = 6, 
         fontcolor = "#3e4989",
         x = 0.75,
         y = 4.875, 
         just = c("left", "top"),
         default.units = "inches")

# Add the samples name
plotText(label = "H3K27me3", 
         fontsize = 6, 
         fontcolor = "#440154",
         x = 0.75,
         y = 5.25, 
         just = c("left", "top"),
         default.units = "inches")

# ChIP-seq signal SCARB2
# Reads BIGWIG file at SCARB2 location - H3K4me1
bw.SCARB2.H3K4me1 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.6a1889bf-7312-4e41-b1cc-e2f3be7a8076.fc.signal.bigwig"),
                               chrom = "chr4",
                               chromstart = 76173617,
                               chromend = 76253817)

# Reads BIGWIG file at SCARB2 location - H3K4me3
bw.SCARB2.H3K4me3 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.4cb9300f-9097-4374-8f2a-4afd8ff855fb.fc.signal.bigwig"),
                               chrom = "chr4",
                               chromstart = 76173617,
                               chromend = 76253817)

# Reads BIGWIG file at SCARB2 location - H3K36me3
bw.SCARB2.H3K36me3 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.6586e624-852f-4cac-83df-2325cee19699.fc.signal.bigwig"),
                               chrom = "chr4",
                               chromstart = 76173617,
                               chromend = 76253817)

# Reads BIGWIG file at SCARB2 location - H3K9me3
bw.SCARB2.H3K9me3 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.e8fe7664-906e-4d87-8f3a-6ce88dbeadc6.fc.signal.bigwig"),
                               chrom = "chr4",
                               chromstart = 76173617,
                               chromend = 76253817)

# Reads BIGWIG file at SCARB2 location - H3K27me3
bw.SCARB2.H3K27me3 = readBigwig(file = paste0(bw.IHEC.path,
                                             "ihec.chipseq.ihec-chipseq-containerv1.1.4.IHECRE00000986.6.5f5024a9-2e3f-4527-82ec-0ed04bf4a848.fc.signal.bigwig"),
                               chrom = "chr4",
                               chromstart = 76173617,
                               chromend = 76253817)

# Add a scale next to the bigwig files - check the maximum to choose
scale.IHEC.SCARB2.max = max(c(bw.SCARB2.H3K4me1$score,
                              bw.SCARB2.H3K4me3$score,
                              bw.SCARB2.H3K36me3$score,
                              bw.SCARB2.H3K9me3$score,
                              bw.SCARB2.H3K27me3$score)) %>% 
  round(.,
        digits = 1.0)

print(scale.IHEC.SCARB2.max)

# Define parameters for the regions
region.p.IHEC.SCARB2 = pgParams(chrom = "chr4",
                         chromstart = 76173717,
                         chromend = 76253717,
                         assembly = "hg38",
                         range = c(0, scale.IHEC.SCARB2.max))

# Add the genomic label
plotGenomeLabel(
  chrom = "chr4",
  chromstart = 76173717,
  chromend = 76253717,
  assembly = "hg38",
  x = 0.75,
  y = 6.5,
  length = 4.0,
  default.units = "inches",
  scale = "Mb")

# ChIPseq signal H3K4me1
H3K4me1_SCARB2 = plotSignal(
  data = bw.SCARB2.H3K4me1, 
  params = region.p.IHEC.SCARB2,
  fill = "#b5de2b", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 6.75, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K4me1_SCARB2, 
          at = c(0, scale.IHEC.SCARB2.max),
          fontsize = 6)

# ChIPseq signal H3K4me3
H3K4me3_SCARB2 = plotSignal(
  data = bw.SCARB2.H3K4me3, 
  params = region.p.IHEC.SCARB2,
  fill = "#35b779", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 7.125, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K4me3_SCARB2, 
          at = c(0, scale.IHEC.SCARB2.max),
          fontsize = 6)

# ChIPseq signal H3K36me3
H3K36me3_SCARB2 = plotSignal(
  data = bw.SCARB2.H3K36me3, 
  params = region.p.IHEC.SCARB2,
  fill = "#26828e", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 7.5, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K36me3_SCARB2, 
          at = c(0, scale.IHEC.SCARB2.max),
          fontsize = 6)

# ChIPseq signal H3K9me3
H3K9me3_SCARB2 = plotSignal(
  data = bw.SCARB2.H3K9me3, 
  params = region.p.IHEC.SCARB2,
  fill = "#3e4989", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 7.875, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K9me3_SCARB2, 
          at = c(0, scale.IHEC.SCARB2.max),
          fontsize = 6)

# ChIPseq signal H3K27me3
H3K27me3_SCARB2 = plotSignal(
  data = bw.SCARB2.H3K27me3, 
  params = region.p.IHEC.SCARB2,
  fill = "#440154", 
  alpha = 0.7, 
  linecolor = NA,
  x = 0.75, 
  y = 8.25, 
  width = 4.0, 
  height = 0.25,
  just = c("left", "top"), 
  default.units = "inches")

annoYaxis(plot = H3K27me3_SCARB2, 
          at = c(0, scale.IHEC.SCARB2.max),
          fontsize = 6)

# Gene track
plotGenes(
  chrom = "chr4",
  chromstart = 76173717,
  chromend = 76253717,
  assembly = assembly(Genome = "hg38refGene", 
                      TxDb = "TxDb.Hsapiens.UCSC.hg38.refGene", 
                      OrgDb = "org.Hs.eg.db"),
  fontcolor = "black",
  fill = "black",
  x = 0.75,
  y = 8.625,
  width = 4.0,
  height = 0.375,
  just = c("left", "top"),
  default.units = "inches")

# Add the samples name
plotText(label = "H3K4me1", 
         fontsize = 6, 
         fontcolor = "#b5de2b",
         x = 0.75,
         y = 6.75, 
         just = c("left", "top"),
         default.units = "inches")

# Add the samples name
plotText(label = "H3K4me3", 
         fontsize = 6, 
         fontcolor = "#35b779",
         x = 0.75,
         y = 7.125, 
         just = c("left", "top"),
         default.units = "inches")

# Add the samples name
plotText(label = "H3K36me3", 
         fontsize = 6, 
         fontcolor = "#26828e",
         x = 0.75,
         y = 7.5, 
         just = c("left", "top"),
         default.units = "inches")

# Add the samples name
plotText(label = "H3K9me3", 
         fontsize = 6, 
         fontcolor = "#3e4989",
         x = 0.75,
         y = 7.875, 
         just = c("left", "top"),
         default.units = "inches")

# Add the samples name
plotText(label = "H3K27me3", 
         fontsize = 6, 
         fontcolor = "#440154",
         x = 0.75,
         y = 8.25, 
         just = c("left", "top"),
         default.units = "inches")

#### PANEL B - 
# text Supplementary Figure 1
# text B
plotText(label = "B", 
         fontsize = 12,
         x = 0.25, 
         y = 8.5, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# Load IHEC RNA-seq data (TPM)
IHEC.exp = read_delim("/home/vagrant/Documents/IHEC/RNA_seq/genes_tpm.csv",
                      delim = ",",
                      col_names = TRUE)

# Genes of interest to be plotted
goi_ens.filt = goi_ens %>% 
  as_tibble() %>% 
  dplyr::filter(ENSEMBL != "ENSG00000274577")

IHEC.exp %>% 
  # Remove the dot and number in Ensembl IDS
  mutate(id_col = gsub("\\..*", 
                       "", 
                       id_col)) %>% 
  dplyr:::filter(id_col %in% goi_ens.filt$ENSEMBL) %>% 
  pivot_longer(cols = starts_with("IHEC"),
               names_to = "Sample",
               values_to = "TPM") %>% 
  mutate(Sample = gsub("",
                       "",
                       Sample))

dev.off()
```

### SUPPLEMENTARY FIGURE 2
Plot genes of interest from the TH REP2 mCHERRY cell line RNAseq data to demonstrate that expression is similar
```{r suppFig2, eval = FALSE, echo = TRUE}
############################################
########## SUPPLEMENTARY FIGURE 2 ##########
############################################

# Save as a PDF
pdf("/home/vagrant/Manuscript_1/SUPPLEMENTARY_FIGURE2/SUPPLEMENTARY_FIGURE2.pdf",
    width = 13, 
    height = 15)

# Create a A4 blank page
pageCreate(width = 8.3, 
           height = 11.7, 
           default.units = "inches",
           showGuides = FALSE)

#### PANEL A - 
# text Supplementary Figure 2
plotText(label = "Supplementary Figure 2", 
         fontsize = 14,
         x = 0.25, 
         y = 0.25, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")
# text A
plotText(label = "A", 
         fontsize = 12,
         x = 0.25, 
         y = 0.5, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# Load data
dat.THREP2.FPKM = read_delim("/home/vagrant/epifunc/RNAseq/TH_REP2_mCHERRY/180604_HT-Rep2_fpkm.tsv",
                      delim = "\t", 
                      col_names = TRUE)

dat.THREP2.FPKM %>% 
  dplyr::filter(gene_name == "SCARB2") %>% 
  summary()

# Filter for genes of interest - BAG3 and LHX1
BAG3.LHX1 = c("BAG3",
              "LHX1")

IDUA.ZBTB14 = c("IDUA",
                "ZBTB14",
                "SLC26A1")

SCARB2.NR2C2 = c("SCARB2",
                 "NR2C2",
                 "FAM47E")

# Reformat the matrix and take only positively sorted neurons
dat.THREP2.FPKM = dat.THREP2.FPKM %>% 
  pivot_longer(starts_with("TH"), 
               names_to = "Samples",
               values_to = "FPKM") %>% 
  dplyr::filter(str_detect(Samples, paste("SmNPC", "pos", sep = "|"))) %>% 
  mutate(Cond = case_when(str_detect(Samples, "D15pos") ~ "Positively sorted neurons D15",
                          str_detect(Samples, "D30pos") ~ "Positively sorted neurons D30",
                          TRUE ~ "smNPCs"),
         Cond = factor(Cond, 
                       levels = c("smNPCs", 
                                  "Positively sorted neurons D15",
                                  "Positively sorted neurons D30")))

# Plot the expression of BAG3 and LHX1
TH.REP2_BAG3.LHX1.exp = ggboxplot(dat.THREP2.FPKM %>% 
                            dplyr::filter(gene_name %in% BAG3.LHX1),
                          x = "Cond", 
                          y = "FPKM", 
                          add = "jitter", 
                          fill = "Cond", 
                          palette = c("#A9A9A9", "#B22222", "#DC143C"),
                          facet.by = "gene_name",
                          xlab = "", 
                          ylab = "Fragments Per Kilobase Million (FPKM)") +
  scale_x_discrete(labels = c("smNPC",
                              expression("mDAN D15"),
                              expression("mDAN D30"))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))

# Place the plot  
plotGG(plot = TH.REP2_BAG3.LHX1.exp, 
       x = 0.5, 
       y = 0.625, 
       width = 2, 
       height = 3.5,
       just = c("left", "top"), 
       default.units = "inches")

# Plot the expression of IDUA and ZBTB14
TH.REP2_IDUA.ZBTB14.exp = ggboxplot(dat.THREP2.FPKM %>% 
                            dplyr::filter(gene_name %in% IDUA.ZBTB14) %>% 
                              mutate(gene_name = factor(gene_name,
                                                        levels = c("IDUA", "ZBTB14", "SLC26A1"))),
                          x = "Cond", 
                          y = "FPKM", 
                          add = "jitter", 
                          fill = "Cond", 
                          palette = c("#A9A9A9", "#B22222", "#DC143C"),
                          facet.by = "gene_name",
                          xlab = "", 
                          ylab = "Fragments Per Kilobase Million (FPKM)") +
  scale_x_discrete(labels = c("smNPC",
                              expression("mDAN D15"),
                              expression("mDAN D30"))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))

# Place the plot  
plotGG(plot = TH.REP2_IDUA.ZBTB14.exp, 
       x = 2.75, 
       y = 0.625, 
       width = 2.5, 
       height = 3.5,
       just = c("left", "top"), 
       default.units = "inches")

# Plot the expression of SCARB2 and NR2C2
TH.REP2_SCARB2.NR2C2.exp = ggboxplot(dat.THREP2.FPKM %>% 
                            dplyr::filter(gene_name %in% SCARB2.NR2C2) %>% 
                              mutate(gene_name = factor(gene_name,
                                                        levels = c("SCARB2", "NR2C2", "FAM47E"))),
                          x = "Cond", 
                          y = "FPKM", 
                          add = "jitter", 
                          fill = "Cond", 
                          palette = c("#A9A9A9", "#B22222", "#DC143C"),
                          facet.by = "gene_name",
                          xlab = "", 
                          ylab = "Fragments Per Kilobase Million (FPKM)") +
  scale_x_discrete(labels = c("smNPC",
                              expression("mDAN D15"),
                              expression("mDAN D30"))) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))

# Place the plot  
plotGG(plot = TH.REP2_SCARB2.NR2C2.exp, 
       x = 5.5, 
       y = 0.625, 
       width = 2.5, 
       height = 3.5,
       just = c("left", "top"), 
       default.units = "inches")

#### PANEL B - 
# text B
plotText(label = "B", 
         fontsize = 12,
         x = 0.25, 
         y = 4.5, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# FOUNDIN-PD data
# Load
FIPD = read_delim("/home/vagrant/epifunc/PPMI_data/FOUNDIN-PD_150.9_RNAB/aggregated_expression/cpmTable.tsv",
                  delim = "\t",
                  col_names = TRUE)

#FIPD = read_delim("/home/vagrant/epifunc/PPMI_data/FOUNDIN-PD_150.9_RNAB/aggregated_expression/countTable.tsv",
#                  delim = "\t",
#                  col_names = TRUE)

# Filter for genes of interest and get the ensembl and gene names
goi_ens = select(org.Hs.eg.db, 
                 keys = c("BAG3",
                          "LHX1",
                          "IDUA",
                          "ZBTB14",
                          "SLC26A1",
                          "SCARB2",
                          "NR2C2",
                          "FAM47E"), 
                 columns = c("SYMBOL",
                             "ENSEMBL"), 
                 keytype = "SYMBOL")

# 2 LHX1 ensembl IDs came back. Check on the Ensembl website which one is correct
goi_ens = goi_ens %>% 
  as_tibble() %>% 
  dplyr::filter(ENSEMBL != "ENSG00000274577") %>% 
  dplyr::rename(ENSEMBL.not.dot = ENSEMBL)

# Load the information abour iPSCs cell lines
info_iPSCs = read_delim("/home/vagrant/epifunc/PPMI_data/Participant_Status_02Jun2023.csv",
                  delim = ",",
                  col_names = TRUE)

info_iPSCs = info_iPSCs %>% 
  mutate(PATNO = as.character(PATNO))

# Filter the matrix of expression for my genes of interest
FIPD.fin = FIPD %>% 
  #mutate(Geneid = str_replace(Geneid, "\\..", "")) %>% 
  dplyr::filter(str_detect(Geneid, paste(goi_ens$ENSEMBL.not.dot, collapse = "|"))) %>% 
  dplyr::rename(ENSEMBL = Geneid) %>% 
  mutate(ENSEMBL.not.dot = gsub("\\..*", "", ENSEMBL)) %>% 
  dplyr::select(ENSEMBL, ENSEMBL.not.dot, everything()) %>% 
  #left_join(goi_ens, by = "ENSEMBL") %>% 
  pivot_longer(cols = starts_with("RNAB"), names_to = "Samples", values_to = "CPM") %>% 
  mutate(PATNO = str_extract(Samples, "[^RNAB_PPMI_]+"),
         Day = str_extract(Samples, "da+\\d+")) %>% 
  left_join(info_iPSCs,
            by = "PATNO") %>% 
  left_join(goi_ens, 
            by = "ENSEMBL.not.dot") %>% 
  dplyr::select(ENSEMBL:Day, COHORT_DEFINITION, SYMBOL) %>% 
  mutate(Condition = case_when(COHORT_DEFINITION == "Healthy Control" ~ "Healthy\ncontrol",
                               COHORT_DEFINITION == "Parkinson's Disease" ~ "Parkinson's\ndisease",
                               TRUE ~ COHORT_DEFINITION)) %>% 
  #group_by(SYMBOL, Day, COHORT_DEFINITION) %>% 
  group_by(ENSEMBL, Day) %>% 
  mutate(AVG_CPM = mean(CPM, na.rm = TRUE))

# Plot expression of all putative target genes and TFs except SCARB2 (will be plot alone)
FIPD.exp.pl = ggboxplot(FIPD.fin %>% 
                          dplyr::filter(SYMBOL != "SCARB2"),
                     x = "Day", 
                     y = "AVG_CPM", 
                     add = "jitter", 
                     fill = "Day", 
                     palette = c("#482677FF", "#2D708EFF", "#29AF7FFF"),
                     facet.by = "SYMBOL",
                     xlab = "", 
                     ylab = "CPM") +
  scale_x_discrete(labels = c("Day 0",
                              "Day 25",
                              "Day 65")) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90),
        strip.text.x = element_text(size = 6))

# Place the plot  
plotGG(plot = FIPD.exp.pl, 
       x = 0.5, 
       y = 4.75 ,
       width = 6, 
       height = 4.5,
       just = c("left", "top"), 
       default.units = "inches")

# Plot expression of SCARB2
FIPD.exp.SCARB2.pl = ggboxplot(FIPD.fin %>% 
                          dplyr::filter(SYMBOL == "SCARB2"),
                     x = "Day", 
                     y = "AVG_CPM", 
                     add = "jitter", 
                     fill = "Day", 
                     palette = c("#482677FF", "#2D708EFF", "#29AF7FFF"),
                     facet.by = "SYMBOL",
                     xlab = "", 
                     ylab = "CPM") +
  scale_x_discrete(labels = c("Day 0",
                              "Day 25",
                              "Day 65")) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90),
        strip.text.x = element_text(size = 6))

# Place the plot  
plotGG(plot = FIPD.exp.SCARB2.pl, 
       x = 0.5, 
       y = 9.75 ,
       width = 6, 
       height = 3.5,
       just = c("left", "top"), 
       default.units = "inches")

dev.off()
```

### SUPPLEMENTARY FIGURE 3
Plot FACS results for experiments where knockdown were performed at day 9 and samples collected 3 and 6 days post-transduction
```{r suupFig3, eval = FALSE, echo = TRUE}
############################################
########## SUPPLEMENTARY FIGURE 3 ##########
############################################

# Save as a PDF
pdf("/home/vagrant/Manuscript_1/SUPPLEMENTARY_FIGURE3/SUPPLEMENTARY_FIGURE3.pdf",
    width = 13, 
    height = 15)

# Create a A4 blank page
pageCreate(width = 8.3, 
           height = 11.7, 
           default.units = "inches",
           showGuides = TRUE)

#### PANEL A - 
# text Supplementary Figure 3
plotText(label = "Supplementary Figure 3", 
         fontsize = 14,
         x = 0.25, 
         y = 0.25, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")
# text A
plotText(label = "A", 
         fontsize = 12,
         x = 0.25, 
         y = 0.5, 
         just = "left", 
         default.units = "inches",
         fontface = "bold")

# Read the facs files (.fcs)
#facs.path = "/home/vagrant/Documents/Wetlab/epifunc/"
facs.path = "/Volumes/deborah.gerard/Documents/Wetlab/epifunc/"

facs_6days_PT_N1.fs = read.flowSet(path = paste0(facs.path,
                                              "2022.11_DG_KD_SCARB2_NR2C2_1/FACS/RAW_DATA/NEW_EXPORT_06122022_1624/2022.11_DG_KD_SCARB2_NR2C2_1/"),
                                pattern = ".fcs",
                               alter.names = TRUE,
                               truncate_max_range = FALSE)

# Check what channels are present
colnames(facs_6days_PT_N1.fs)

# Check sample
pData(facs_6days_PT_N1.fs)

# Rename some for clarity
colnames(facs_6days_PT_N1.fs)[colnames(facs_6days_PT_N1.fs) == "FITC.A"] = "GFP"
colnames(facs_6days_PT_N1.fs)[colnames(facs_6days_PT_N1.fs) == "mCherry..A"] = "mCHERRY"
colnames(facs_6days_PT_N1.fs)[colnames(facs_6days_PT_N1.fs) == "SYTOX_BLUE..A"] = "SYTOX BLUE"

# To be able to add gates, the object must be transformed as a GatingSet object
facs_6days_PT_N1.gs = GatingSet(facs_6days_PT_N1.fs)

# Define neurons gate
dead.neur.gate = polygonGate(filterId = "Neurons",
                        "FSC.A" = c(5e4, 2e5, 2e5, 5e4),
                        "SSC.A" = c(0e4, 125e3, 2.e5, 1e5))

# Plot SYTOX BLUE positive CTRL
ggcyto(facs_6days_PT_N1.gs[[1]],
       aes(x = FSC.A, 
           y = SSC.A),
       subset = "root") + 
  geom_hex(bins = 200) +
  geom_gate(dead.neur.gate) +
  ggcyto_par_set(limits = "instrument")

# The gate for dead neurons has been defined -> add the gate to the gateset and recompute
add(facs_6days_PT_N1.gs,
    dead.neur.gate)

recompute(facs_6days_PT_N1.gs)

dev.off()
```

SCARB2 snp (rs1465922) for samples with whole genome sequencing data (75 cases)
```{r, eval= FALSE, echo = TRUE}
# Load data
SCARB2.SNP.WGS.75 = read_excel("/Volumes/deborah.gerard/Documents/epifunc/Final.WGS.SCARB2.OK_from_Sinthu.xlsx")

# Filter for rs1465922
SCARB2.SNP.WGS.75 %>% 
  filter(avsnp150 == "rs1465922") %>% 
  select(samples) %>% 
  separate(samples,
           sep = ", ",
           into = )

```

```{r}
sessionInfo()
```